---
title: "peripheral blood immune cell type signatures: comparing cell type between control and aging datasets"
author: "Tanya Karagiannis"
date: "04/19/20"
output: html_document
---


Purpose: 
Comparison of subpopulations between natgen2018 data and pnas2019 aging data
Form consensus cell types

Load libraries
```{r}
library(dplyr)
library(Seurat)
library(useful)
library(cowplot)

library(harmony, lib.loc = "/restricted/projectnb/montilab-p/R_pkgs/CentOS7/R_4.0.0/")
#library(devtools)
#withr::with_libpaths("/restricted/projectnb/montilab-p/R_pkgs/CentOS7/R_3.6.0",install_github("immunogenomics/harmony"))

PBMC_SC <- readRDS(file = "/restricted/projectnb/montilab-p/projects/longevity/scCentenarian/Public_Reanalysis/pnas2019/data/pnas2019.rds")
PBMC_45 <- readRDS(file = "/restricted/projectnb/montilab-p/projects/longevity/scCentenarian/Public_Reanalysis/natgen2018/data/natgen2018.rds")

```

Change format of names in pnas2019 to match format in natgen2018
```{r}
PBMC_SC.dat <- GetAssay(PBMC_SC)
PBMC_SC.counts <- PBMC_SC.dat@counts
PBMC_SC.log <- PBMC_SC.dat@data
rownames(PBMC_SC.counts) <- make.names(rownames(PBMC_SC.counts))
rownames(PBMC_SC.log) <- make.names(rownames(PBMC_SC.log))

PBMC_SC[['RNA']]@counts <- PBMC_SC.counts
PBMC_SC[['RNA']]@data <- PBMC_SC.log
```

Subpopulations visualized for each dataset based on separate analyses
```{r}
# Projecting identities TSNE visualization
p1 <- DimPlot(PBMC_45, reduction = "umap", group.by = "subpop", label = TRUE)
p2 <- DimPlot(PBMC_SC, reduction = "umap", group.by = "subpop", label = TRUE)
plot_grid(p1, p2)
```


Merge datasets in one seurat object and run PCA 
```{r}
pbmc <- merge(PBMC_45, PBMC_SC, add.cell.ids = c("NATGEN", "PNAS"), project = "PBMC.IMMUNE") %>%
    Seurat::NormalizeData(verbose = FALSE) %>%
    FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
    ScaleData(verbose = FALSE) %>% 
    RunPCA(pc.genes = pbmc@var.genes, npcs = 20, verbose = FALSE)

head(pbmc@meta.data)

#add variable batch to distinguish datasets
pbmc@meta.data$batch <- c(rep("NATGEN", ncol(PBMC_45)), rep("PNAS", ncol(PBMC_SC)))
pbmc <- ScaleData(pbmc, split.by = "batch")

#Visualize PCA
p1 <- DimPlot(object = pbmc, reduction = "pca", pt.size = .1, group.by = "batch")
p2 <- VlnPlot(object = pbmc, features = "PC_1", group.by = "batch", pt.size = .1)
plot_grid(p1,p2)

```

Function to calculate average silhoutte score
```{r}
batch_sil <- function(pca.data, batch, nPCs = 20){
  require(cluster)
    # in scone, they use svd to compute principal components.
    # For now, we'll keep the PCA object created in
    # prcomp to be consistent
    # proj <- svd(scale(t(expr),center = TRUE,scale = TRUE),
    #             nu = 3, nv =0)$u

    dd <- as.matrix(dist(pca.data[, seq_len(nPCs)]))
    summary(silhouette(as.numeric(batch), dd))$avg.width
}
```

Calculate average silhoutte score for each cell type population before batch correction with Harmony
```{r}
#local silhouette score
cell.types <- levels(as.factor(pbmc$ct.consensus))
table(pbmc$ct.consensus, pbmc$batch)
cell.types <- cell.types[-which(cell.types %in% c("T gamma delta","CD4 T Cytotoxic"))]
#pca.data <- pbmc@reductions$pca@cell.embeddings

pca.sil.score <- lapply(1:length(cell.types), function(x){
  ct.order <- which(pbmc@meta.data$ct.consensus == cell.types[x])
  ct.cells <- rownames(pbmc@meta.data)[ct.order]
  pca.data <- pbmc@reductions$pca@cell.embeddings[ct.cells,]
  batch <- as.factor(pbmc@meta.data$batch[ct.order])

  pca.sil <- batch_sil(pca.data, batch, nPCs = 20)
  return(pca.sil)
})
pca.sil.score <- unlist(pca.sil.score)
names(pca.sil.score) <- cell.types

#local silhouette score
cell.types <- levels(as.factor(pbmc$ct.consensus))
table(pbmc$ct.consensus, pbmc$batch)
cell.types <- cell.types[-which(cell.types %in% c("T gamma delta","CD4 T Cytotoxic"))]
#pca.data <- pbmc@reductions$pca@cell.embeddings

```

Run Harmony to correct for batch differences in PCA space
```{r}
pbmc <- pbmc %>% 
    RunHarmony("batch", plot_convergence = TRUE)

saveRDS(pbmc, file = "/restricted/projectnb/montilab-p/projects/longevity/scCentenarian/Public_Reanalysis/Data/pbmc.rds")

p1 <- DimPlot(object = pbmc, reduction = "harmony", pt.size = .1, group.by = "batch")
p2 <- VlnPlot(object = pbmc, features = "harmony_1", group.by = "batch", pt.size = .1)
plot_grid(p1,p2)

```


Average silhoutte score for each cell type after batch correction
```{r}
#local average silhoutte score
harmony.sil.score <- lapply(1:length(cell.types), function(x){
  ct.order <- which(pbmc@meta.data$ct.consensus == cell.types[x])
  ct.cells <- rownames(pbmc@meta.data)[ct.order]
  harmony.data <- pbmc@reductions$harmony@cell.embeddings[ct.cells,]
  batch <- as.factor(pbmc@meta.data$batch[ct.order])
  harmony.sil <- batch_sil(harmony.data, batch, nPCs = 20)
  return(harmony.sil)
})
harmony.sil.score <- unlist(harmony.sil.score)
names(harmony.sil.score) <- cell.types
sil.score <- cbind.data.frame("pre-correction" = pca.sil.score, "post-correction" = harmony.sil.score)
write.csv(sil.score, "public.sil.score.csv")
```

Run UMAP based on top 20 Harmony corrected PCs
```{r}
pbmc <- pbmc %>% 
    RunUMAP(reduction = "harmony", dims = 1:20) %>%
    FindNeighbors(reduction = "harmony", dims = 1:20) %>% 
    FindClusters(resolution = 0.8) %>% 
    identity()

p1 <- DimPlot(object = pbmc, reduction = "umap", pt.size = .1, group.by = "seurat_clusters", label = TRUE)
p2 <- DimPlot(object = pbmc, reduction = "umap", pt.size = .1, group.by = "subpop", label = T)
plot_grid(p1,p2)

```

Feature average expression across separately identified subpopulations
```{r}

Idents(pbmc) <- "subpop"
DotPlot(pbmc, features = rev(c("CD3D", "CD3G", "TRAC", #tcells
                               "CCR7","S100A4","CREM", #naive vs activated
                              "CD8A","CD8B","TRDC",#cd8 and tdeltagamma
                              "GZMB","KLRF1","GNLY", "NKG7", #nk cells
                              "CD19", "CD79A", "MS4A1", #bcells
                              "CD14","FCGR3A","S100A8", "S100A12","CD163", #monocytes
                              "ITGB3","CMTM5","PPBP", "GNG11", #megakaryocytes
                              "MALAT1", 
                              "HLA.DQA1", "LILRA4","TSPAN13", "GPR183", "IL3RA", #DC
                              "IGJ", "MZB1", #plasma cells
                              "HBA1","HBD1", "HBA2", "HBB", #erythrocytes
                              "MKI67","CDK1" )) #MKI 
) + RotatedAxis()
```


Set consensus populations
```{r}

Idents(pbmc) <- "seurat_clusters"
B.naive <- WhichCells(pbmc, idents=8)
B.memory <- WhichCells(pbmc, idents=c(12))
EC.cells <- WhichCells(pbmc, idents=c(9,16,22,24))
MKI.cells <- WhichCells(pbmc, idents=c(21))


Idents(pbmc) <- "subpop"
Idents(pbmc, cells=B.naive) <- "BC Naive"
Idents(pbmc, cells=B.memory) <- "BC Memory"
Idents(pbmc, cells=EC.cells) <- "EC"
Idents(pbmc, cells=MKI.cells) <- "MKI"
Idents(pbmc, cells=WhichCells(pbmc, idents=c("Pro.myeloid"))) <- "MGK"
Idents(pbmc, cells=WhichCells(pbmc, idents=c("CD8 T V2"))) <- "BC Naive"
Idents(pbmc, cells=WhichCells(pbmc, idents=c("HSC"))) <- "pDC"

pbmc$ct.consensus <- Idents(pbmc)

saveRDS(pbmc, file = "/restricted/projectnb/montilab-p/projects/longevity/scCentenarian/Public_Reanalysis/Data/pbmc_public.rds")

Idents(pbmc) <- "group"
CT.cells <- WhichCells(pbmc, idents="CT")
EL.cells <- WhichCells(pbmc, idents = "SC")

Idents(pbmc) <- "age.group"
Idents(pbmc, cells=CT.cells) <- "CT"
Idents(pbmc, cells=EL.cells) <- "EL"

pbmc$age.group <- Idents(pbmc)

Idents(pbmc) <- "sample.ID"
Female.cells <- WhichCells(pbmc, idents = c("SC1","SC3","SC4","SC5","SC6","CT2","CT3","CT4"))
Male.cells <- WhichCells(pbmc, idents = c("SC2","SC7","CT1","CT5"))
Idents(pbmc) <- "Sex"
Idents(pbmc, cells=Female.cells) <- "Female"
Idents(pbmc, cells=Male.cells) <- "Male"
pbmc$sex <- Idents(pbmc)

p1 <- DimPlot(object = pbmc, reduction = "umap", pt.size = .1, group.by = "ct.consensus", label = TRUE)
p2 <- DimPlot(object = pbmc, reduction = "umap", pt.size = .1, group.by = "ct.consensus",split.by = "batch", label = T )
plot_grid(p1,p2)

```

Feature average expression of consensus cell types
```{r}
pbmc$ct.consensus <- factor(pbmc$ct.consensus,levels = rev(c("CD4 T NonCytotoxic","CD4 T Cytotoxic","CD8 T Cytotoxic","T gamma delta","NK","BC Naive","BC Memory","M14","M16","MGK","mDC","pDC","Plasma","EC","MKI")))

Idents(pbmc) <- "ct.consensus"
DotPlot(pbmc, features = rev(c("CD3D", "CD3G", "TRAC", #tcells
                               "CCR7","S100A4","CREM", #naive vs activated
                              "CD8A","CD8B","TRDC",#cd8 and tdeltagamma
                              "GZMB","KLRF1","GNLY", "NKG7", #nk cells
                              "CD19", "CD79A", "MS4A1", #bcells
                              "CD14","FCGR3A","S100A8", "S100A12","CD163", #monocytes
                              "ITGB3","CMTM5","PPBP", "GNG11", #megakaryocytes
                              "MALAT1", 
                              "HLA.DQA1", "LILRA4","TSPAN13", "GPR183", "IL3RA", #DC
                              "IGJ", "MZB1", #plasma cells
                              "HBA1","HBD1", "HBA2", "HBB", #erythrocytes
                              "MKI67","CDK1" )) #MKI 
) + RotatedAxis()

```

Average scaled expression module scores for immune cell type signatures
```{r}
vlnplot <- function(seurat_object, metadata, group, cell.type = NULL, color.box = color.box){
  
  #seurat_object <- subset(seurat_object, idents = cell.type)
  
  
  #Res <- seurat_object@meta.data[, c("ct.consensus", names(genesets))]
  Res <- cbind.data.frame(group, metadata)
  colnames(Res)[1] <- "clusters"
  require(reshape2)
  ResFlat <- reshape2::melt(Res)
  ResFlat <- ResFlat %>% filter(clusters == cell.type)
  #color.box <- rep("grey", length(genesets))
  #color.box[cell.type %in% names(genesets)] <- "blue"
  
  require(ggplot2)
  ggplot2::ggplot(ResFlat, aes(
    x = variable, y = value)) +
    #ggplot2::geom_violin() +
    ggplot2::geom_boxplot() +
    #ggplot2::geom_jitter() +
    ggplot2::facet_wrap(~clusters) +
    theme_classic(base_size = 30) +
    #facet_grid(clusters~variable)+
    aes(fill = as.factor(variable)) +
  scale_color_manual(values = color.box)+
    theme(legend.position = "none")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    labs(x = "genesets", y = "absolute average module score")
  
}

```

Average expression of geneset signatures
```{r}
#Main populations
Idents(pbmc_public) <- "ct.consensus"
Idents(pbmc_public, cells = WhichCells(pbmc_public, idents = c("BC Naive", "BC Memory", "Plasma"))) <- "B.cells"
Idents(pbmc_public, cells = WhichCells(pbmc_public, idents = c("CD4 T Naive", "CD4 T Memory","CD4 T Cytotoxic", "CD8 T","T gamma delta"))) <- "T.cells"
Idents(pbmc_public, cells = WhichCells(pbmc_public, idents = c("M14", "M16","mDC", "pDC","MGK","EC"))) <- "Myeloid.cells"
pbmc_public$main.ct <- Idents(pbmc_public)

#Biocarta = msigdbr(species = "Homo sapiens", category = "C2", subcategory = "CP:BIOCARTA")
#msigdbr_list = split(x =Biocarta$gene_symbol, f = Biocarta$gs_name)
#Biocarta <- msigdbr_list[c("BIOCARTA_THELPER_PATHWAY", "BIOCARTA_TCYTOTOXIC_PATHWAY")]
#Biocarta <- list(T.cell = unique(c(Biocarta[["BIOCARTA_THELPER_PATHWAY"]],Biocarta[["BIOCARTA_TCYTOTOXIC_PATHWAY"]])))

pbmc_public1 <- AddModuleScore(pbmc_public, features = immune.genesets[c("B.cell", "Early.lymphoid.T.lymphocyte", "NK cells resting", "Monocytes", "DC.precursor")], name = c("B.cell", "Early.lymphoid.T.lymphocyte", "NK cells resting", "Monocytes", "DC.precursor"), search = T)

cell.types <- levels(droplevels(pbmc_public1$main.ct))
plot.sets <- lapply(1:length(cell.types), function(x){
  Idents(pbmc_public1) <- "main.ct"
  pbmc_public1 <- subset(pbmc_public1, idents = c(cell.types[x]))
  group <- pbmc_public1$main.ct
  meta.data <- pbmc_public1@meta.data[,c(16:20)]
  #meta.data <- abs(meta.data)
  #meta.data <- avgGeneExp(Tcell1, CIBERSORT_genesets)
  plot.sets <- vlnplot(pbmc_public1, meta.data,group = group, cell.type = cell.types[x]) 
  
  return(plot.sets)
})

#B cell subpopulations
Idents(pbmc_public) <- "ct.consensus"
Bcells <- subset(pbmc_public, idents = c("BC Naive", "BC Memory", "Plasma"))
Bcells1 <- AddModuleScore(Bcells, features = immune.genesets[c("B cells naive", "B cells memory", "Plasma cells")], name = c("B cells naive", "B cells memory", "Plasma cells"), search = T)


cell.types <- levels(droplevels(Bcells1$ct.consensus))
plot.sets <- lapply(1:length(cell.types), function(x){
  Idents(Bcells1) <- "ct.consensus"
  Bcells1 <- subset(Bcells1, idents = c(cell.types[x]))
  group <- Bcells1$ct.consensus
  meta.data <- Bcells1@meta.data[c(15:17)]
  meta.data <- abs(meta.data)
  #meta.data <- avgGeneExp(Bcells1, CIBERSORT_genesets)
  plot.sets <- vlnplot(Bcells1, meta.data,group = group, cell.type = cell.types[x]) 
  
  return(plot.sets)
})

patchwork::wrap_plots(plotlist = plot.sets, ncol = 1)

#T cell subpopulations
Idents(pbmc_public) <- "ct.consensus"
Tcells <- subset(pbmc_public, idents = c("CD4 T Naive", "CD4 T Memory", "CD8 T", "CD4 T Cytotoxic", "T gamma delta", "BC Naive", "BC Memory", "Plasma","NK"))

CD4Memory.sig <- list("T cells CD4 memory" = unique(do.call(c,immune.genesets[c("T cells CD4 memory resting", "T cells CD4 memory activated")])))
NK.sig <- list("NK cells" = unique(do.call(c,immune.genesets[c("NK cells resting", "NK cells activated")])))

Tcells1 <- AddModuleScore(Tcells, features = c(immune.genesets[c("T cells CD4 naive")], CD4Memory.sig, immune.genesets[c("T cells CD8", "T cells gamma delta", "B cells naive", "B cells memory", "Plasma cells")], NK.sig), name = c(c("T cells CD4 naive", "T cells CD4 memory resting", "T cells CD8", "T cells gamma delta", "B cells naive","B cells memory","Plasma cells","NK cells")), search = T)

cell.types <- levels(droplevels(Tcells1$ct.consensus))
plot.sets <- lapply(1:length(cell.types), function(x){
  Idents(Tcells1) <- "ct.consensus"
  Tcells1 <- subset(Tcells1, idents = c(cell.types[x]))
  group <- Tcells1$ct.consensus
  meta.data <- Tcells1@meta.data[c(15:22)]
  meta.data <- abs(meta.data)
  #meta.data <- avgGeneExp(Tcell1, CIBERSORT_genesets)
  plot.sets <- vlnplot(Tcells1, meta.data,group = group, cell.type = cell.types[x]) 
  
  return(plot.sets)
})

#Myeloid subpopulations
Idents(pbmc_public) <- "ct.consensus"
Myeloid <- subset(pbmc_public, idents = c("M14","M16", "mDC", "pDC", "MGK", "EC"))
Myeloid1 <- AddModuleScore(Myeloid, features = c(immune.genesets[c("Monocyte", "Mono.Mac")] ,Liver.Immune[c("DC1")],immune.genesets[c("pDC.precursor","Megakaryocyte", "Late.Erythroid")]), name = c(c("Monocyte", "Mono.Mac", "mDC","pDC.precursor", "Megakaryocyte", "Late.Erythroid")), search = T)

cell.types <- levels(droplevels(Myeloid1$ct.consensus))
plot.sets <- lapply(1:length(cell.types), function(x){
  Idents(Myeloid1) <- "ct.consensus"
  Myeloid1 <- subset(Myeloid1, idents = c(cell.types[x]))
  group <- Myeloid1$ct.consensus
  meta.data <- Myeloid1@meta.data[c(15:20)]
  meta.data <- abs(meta.data)
  #meta.data <- avgGeneExp(Tcell1, CIBERSORT_genesets)
  plot.sets <- vlnplot(Myeloid1, meta.data,group = group, cell.type = cell.types[x]) 
  
  return(plot.sets)
})


```



Add back to separate datasets
```{r}
Idents(pbmc) <- "batch"
natgen <- subset(pbmc, ident = "NATGEN")
pnas <- subset(pbmc, ident = "PNAS")

PBMC_45 <- AddMetaData(PBMC_45, metadata = natgen@meta.data$ct.consensus, col.name = "ct.consensus")
PBMC_45 <- AddMetaData(PBMC_45, metadata = natgen@meta.data$age.group, col.name = "age.group")
saveRDS(PBMC_45,"/restricted/projectnb/montilab-p/projects/longevity/scCentenarian/Public_Reanalysis/natgen2018/data/natgen2018.rds")
PBMC_SC@meta.data$ct.consensus <- pnas$ct.consensus
PBMC_SC@meta.data$age.group <- pnas$age.group
saveRDS(PBMC_SC,"/restricted/projectnb/montilab-p/projects/longevity/scCentenarian/Public_Reanalysis/pnas2019/data/pnas2019.rds")

```
