---
title: "Differential Analysis of cell type specific EL and Aging signatures"
author: "Tanya Karagiannis, Stefano Monti, Paola Sebastiani"
output:
  html_document:
    theme: united
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapse: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
## Functions to run to make nice figures!

#ink colors from anfed/risoink

#' Raw riso ink colors
#' 
#' @keywords internal
.riso_colors <- function() {
  c(
    "black"="#000000",
    "burgundy"="#914E72",
    "blue"="#0078BF",
    "green"="#00A95C",
    "medium-blue"="#3255A4",
    "bright-red"="#F15060",
    "risofederal-blue"="#3D5588",
    "purple"="#765BA7",
    "teal"="#00838A",
    "flat-gold"="#BB8B41",
    "hunter-green"="#407060",
    "red"="#FF665E",
    "brown"="#925F52",
    "yellow"="#FFE800",
    "marine-red"="#D2515E",
    "orange"="#FF6C2F",
    "fluorescent-pink"="#FF48B0",
    "light-gray"="#88898A",
    "metallic-gold"="#AC936E",
    "crimson"="#E45D50",
    "fluorescent-orange"="#FF7477",
    "cornflower"="#62A8E5",
    "sky-blue"="#4982CF",
    "sea-blue"="#0074A2",
    "lake"="#235BA8",
    "indigo"="#484D7A",
    "midnight"="#435060",
    "mist"="#D5E4C0",
    "granite"="#A5AAA8",
    "charcoal"="#70747C",
    "smoky-teal"="#5F8289",
    "steel"="#375E77",
    "slate"="#5E695E",
    "turquoise"="#00AA93",
    "emerald"="#19975D",
    "grass"="#397E58",
    "forest"="#516E5A",
    "spruce"="#4A635D",
    "moss"="#68724D",
    "sea-foam"="#62C2B1",
    "kelly-green"="#67B346",
    "light-teal"="#009DA5",
    "ivy"="#169B62",
    "pine"="#237E74",
    "lagoon"="#2F6165",
    "violet"="#9D7AD2",
    "orchid"="#AA60BF",
    "plum"="#845991",
    "raisin"="#775D7A",
    "grape"="#6C5D80",
    "scarlet"="#F65058",
    "tomato"="#D2515E",
    "cranberry"="#D1517A",
    "maroon"="#9E4C6E",
    "raspberry-red"="#D1517A",
    "brick"="#A75154",
    "light-lime"="#E3ED55",
    "sunflower"="#FFB511",
    "melon"="#FFAE3B",
    "apricot"="#F6A04D",
    "paprika"="#EE7F4B",
    "pumpkin"="#FF6F4C",
    "bright-olive-green"="#B49F29",
    "bright-gold"="#BA8032",
    "copper"="#BD6439",
    "mahogany"="#8E595A",
    "bisque"="#F2CDCF",
    "bubble-gum"="#F984CA",
    "light-mauve"="#E6B5C9",
    "dark-mauve"="#BD8CA6",
    "wine"="#914E72",
    "gray"="#928D88",
    "white"="#FFFFFF",
    "aqua"="#5EC8E5",
    "mint"="#82D8D5",
    "fluorescent-yellow"="#FFE900",
    "fluorescent-red"="#FF4C65",
    "fluorescent-green"="#44D62C"
  )
}

#' Adjust transparency of a hex string
#'
#' @param hex A 6-character hex string (e.g. #000000)
#' @param percent Transparency level from 0-1
#' @return A hex string
#' 
#' @export
hex_transparacy <- function(hex, percent=1) {
  if (percent < 0) percent <- 0
  if (percent > 1) percent <- 1
  percent <- toupper(as.hexmode(floor(percent * 255)))
  percent <- sprintf("%02s", percent)
  hex <- paste0(hex, percent)
  return(hex)
}

#' Riso color names
#'
#' @param sorted Optionally sort values
#' @return A character vector of names
#' 
#' @export
riso_names <- function(sorted=FALSE) {
  if (sorted) {
    return(sort(names(.riso_colors())))   
  } else {
    return(names(.riso_colors()))
  }
}

#' Get a risk ink color
#'
#' @param color Color name
#' @param alpha Color transparency
#' @return A hex string
#' 
#' @export
riso_color <- function(color, alpha=1) {
  riso_colors <- .riso_colors()
  if (color %in% names(riso_colors)) {
    hex <- riso_colors[[color]]
    if (alpha >= 1) {
      return(hex)
    } else {
      return(hex_transparacy(hex, alpha))
    }
  } else {
    cat("Riso Ink Colors: \n")
    print(riso_names(sorted=TRUE))
    stop("Invalid color")
  }
}

#' Plot one or more colors
#'
#' @param colors Character vector of hex strings
#' @param title Optional plot title
#' @return A `gg` object
#' 
#' @import ggplot2
#' @import ggforce 
#' 
#' @export
show_colors <- function(colors, title="") {
  df <- data.frame(hex=colors, y=1, x=seq(length(colors)))
  fill_values <- colors
  names(fill_values) <- colors
  ggplot(df, aes(x0=x, y0=y, r=0.4, fill=hex, expand=1)) + 
    geom_circle() + 
    scale_fill_manual(values=fill_values) +
    coord_equal() + 
    theme_void() +
    labs(title=paste0("  ", title, "\n")) + 
    xlim(0, max(7, nrow(df))+1) +
    theme(legend.position="none")
}


#Plot using ggstyle to easily change specific aspects
#from montilab/best-practices
#function to modify ggplot parameters from montilab/best-practices
ggstyle <- function(font="Helvetica", scale=1) {
  fs <- function(x) x*scale # Dynamic font scaling
  ggplot2::theme(
    plot.title = ggplot2::element_text(family=font, size=fs(26), face="bold", color="#222222"),
    plot.subtitle = ggplot2::element_text(family=font, size=fs(18), margin=ggplot2::margin(0,0,5,0)),
    plot.caption = ggplot2::element_blank(),
    legend.position = "right",
    legend.text.align = 0,
    legend.background = ggplot2::element_blank(),
    legend.title = ggplot2::element_blank(),
    legend.key = ggplot2::element_blank(),
    legend.text = ggplot2::element_text(family=font, size=fs(18), color="#222222"),
    axis.title =  ggplot2::element_text(family=font, size=fs(18), color="#222222"),
    axis.text = ggplot2::element_text(family=font, size=fs(18), color="#222222"),
    axis.text.x = ggplot2::element_text(margin=ggplot2::margin(5, b=10)),
    #axis.ticks = ggplot2::element_blank(),
    axis.line = ggplot2::element_line(color="#222222"),
    panel.grid.minor = ggplot2::element_blank(),
    panel.grid.major.y = ggplot2::element_blank(),
    panel.grid.major.x = ggplot2::element_blank(),
    panel.background = ggplot2::element_blank(),
    strip.background = ggplot2::element_rect(fill="white"),
    strip.text = ggplot2::element_text(size=fs(22), hjust=0)
  )
}

```

## Load libraries and set work directory
```{r, warning = F, message=F}
library(tidyverse)
library(ggplot2)
#library(ggVennDiagram)
#library(ggvenn)
library(stringr)
library(readxl)
library(xlsx)
library(gridExtra)
library(dplyr)
library(Seurat)

work.dir <- "/restricted/projectnb/uh2-sebas/analysis/scCentenarians/"

cell.types <- c("CD4 T Naive", "CD4 T Memory","CD8 T","NK","BC Naive","BC Memory","M14","M16","mDC")

pbmc.combined <- readRDS(file = paste0(work.dir,"Integrated_Data/pbmc_combined.rds"))

```

## Differential genes of aging and EL
### Load rjags results 
```{r}
rjags.res <- lapply(1:length(cell.types), function(i){
  doc.name <- str_replace_all(cell.types[i], fixed(" "), "_")
  DE.lifespan <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  return(DE.lifespan)
})
names(rjags.res) <- cell.types
```

### Select genes with fdr <= 0.05 and min 10% change 
```{r}
#cell type specific significant genes with log fold change, p-value, and fdr 
DE.sig <- rjags.res %>% purrr::map(function(x) as_tibble(x, rownames = "gene")) %>%
              purrr::map(function(x) select(x, gene, MY_Mean, MY_pval, MY_fdr, OY_Mean, OY_pval, OY_fdr, EL_Mean, EL_pval, EL_fdr)) %>%
              purrr::map(function(x) filter(x, MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05)) %>%
              purrr::map(function(x) filter(x, abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1))) 

#add cell type names for each dataframe
DE.sig <- lapply(1:length(DE.sig), function(x){
  res <- DE.sig[[x]]
  res$cell.type <- rep(names(rjags.res)[x], nrow(res))
  return(res)
  })

#bind dataframes together
DE.sig.df <- DE.sig %>% bind_rows %>% 
                        relocate(cell.type, .before = gene) %>%
                        rename("MY_logFC" = MY_Mean, "OY_logFC" = OY_Mean, "EL_logFC" = EL_Mean)

#save cell type specific significant signatures to csv file
#readr::write_csv(DE.sig.df, paste0(work.dir, "Differential_Analysis/DE.gene.sig.csv"))
```

### Boxplot of each cell type specific DE genes across age
```{r, fig.dim = c(20,20)}
DE.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.EL <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.EL$gene <- rownames(DE.EL)
  DE.EL <- DE.EL %>% filter(MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05) %>% filter(abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1)) %>% arrange(EL_fdr)
  
  return(DE.EL$gene)  
})

plot.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  pbmc.set <- Seurat::AverageExpression(pbmc.set, features = DE.sig.genes[[x]], group.by = "sample.ID", slot = "data")
  
  pbmc.set <- reshape2::melt(pbmc.set$RNA)
  colnames(pbmc.set)[1:2] <- c("gene", "sample.ID")
 
  Age <- unlist(sapply(1:nrow(pbmc.set), function(x){unique(pbmc.combined@meta.data$age.quartiles[pbmc.combined@meta.data$sample.ID %in% pbmc.set$sample.ID[x]])}))

pbmc.set$age <- Age
pbmc.set$age <- factor(pbmc.set$age, levels = c("20-39","40-59","60-89", "Extreme Longevity"))
levels(pbmc.set$age) <- c("Younger", "Middle", "Older", "EL")

plot.sig <- ggplot(pbmc.set, aes(age,value)) + geom_boxplot(aes(col = age)) + geom_point() + facet_wrap(~ gene, scale = "free") + theme_bw(base_size = 30) + ggplot2::scale_color_manual(values = c("blue2", "green","red", "maroon")) + ggtitle(doc.name) + ylab("average expression level")


return(plot.sig) 
  
})

plot.sig.genes


```

### Boxplot of each cell type specific DE genes across age: Figure 4C, 4D, 4E
```{r, fig.dim = c(20,20)}
cell.types <- c("CD4 T Naive", "CD4 T Memory", "CD8 T", "BC Naive", "BC Memory", "M16")

DE.sig.genes <- list("CD4 T Naive" = c("STK17A", "LRRN3", "LEF1", "S100A4", "SESN3"),
                     "CD4 T Memory" = c("HLA.DPA1"),
                     "CD8 T" = c("HLA.DPB1"),
                     "BC Naive" = c("NOP53","MAP1LC3B"),
                     "BC Memory" = c("COX16"),
                     "M16" = c("CTSH", "ELOB")
)
  


plot.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.gene <- DE.sig.genes[[x]]
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  pbmc.set <- Seurat::AverageExpression(pbmc.set, features = DE.gene, group.by = "sample.ID", slot = "data", assays = "RNA")
  
  pbmc.set <- tibble::as_tibble(pbmc.set$RNA, rownames = "gene") %>% mutate(gene.name = DE.gene) %>% gather(sample.ID, avg.exprs, BU10:SCN8)
  
  #pbmc.set <- reshape2::melt(pbmc.set$RNA)
  #colnames(pbmc.set)[1:2] <- c("gene", "sample.ID")
 
  Age <- unlist(sapply(1:nrow(pbmc.set), function(x){unique(pbmc.combined@meta.data$age.quartiles[pbmc.combined@meta.data$sample.ID %in% pbmc.set$sample.ID[x]])}))

pbmc.set$age <- Age
pbmc.set$age <- factor(pbmc.set$age, levels = c("20-39","40-59","60-89", "Extreme Longevity"))
levels(pbmc.set$age) <- c("Younger", "Middle", "Older", "EL")

plot.sig <- sapply(1:length(DE.gene), function(y){
  plot.DE <- ggplot(pbmc.set %>% filter(gene.name == DE.gene[y]), aes(age,avg.exprs)) + geom_boxplot(aes(col = age)) + geom_point() + facet_wrap(~ gene.name, scale = "free") + theme_bw(base_size = 20) + ggplot2::scale_color_manual(values = c("blue2", "green","red", "maroon")) + ggtitle(doc.name) + ylab("average expression level")
  
#   ggplot2::ggsave(
#     filename=paste0("Figure_04_", DE.gene[y], ".pdf"),
#     plot=plot.DE,
#     path=paste0(work.dir,"/Figures"),
#     scale=1,
#     width = 8,
#     height=5,
#     units="in",
#     dpi=300
# )
  
})

  
})



```

### Heatmap of cell type specific DE signatures
```{r, eval = F}
lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  pbmc.set <- Seurat::AverageExpression(pbmc.set, features = DE.sig.genes[[x]], group.by = "sample.ID", slot = "data", assays = "RNA")
  
  Age <- unlist(sapply(1:ncol(pbmc.set$RNA), function(y){unique(pbmc.combined@meta.data$age.quartiles[pbmc.combined@meta.data$sample.ID %in% colnames(pbmc.set$RNA)[y]])}))

Age <- factor(Age, levels = c("20-39","40-59","60-89", "Extreme Longevity"))
levels(Age) <- c("Younger", "Middle", "Older", "EL")


require(ComplexHeatmap)
require(circlize)

ha = HeatmapAnnotation(
    age = Age,
    col = list(age = c("Younger" = "blue2", "Middle" = "green", "Older" = "red", "EL" = "maroon")
    )
)

pbmc.set_scaled = na.omit(t(apply(pbmc.set$RNA, 1, scale)))

h1 <- ComplexHeatmap::Heatmap(pbmc.set_scaled, cluster_rows = F, cluster_columns = T, name = "scaled average expression", show_row_names = T, show_column_names = F, column_title = doc.name, row_names_gp = gpar(fontsize = 12), top_annotation = ha)

return(h1) 
  
})



```
