---
title: "Cell type specific composition analysis across age groups of the human lifespan: Multinomial Logistic Regression"
output: html_document
---

Summary:

- Apply multinomial logistic regression model (RJAGS) to cell counts of samples across cell types 
- Output: Cell type specific probability estimates of composition per age group/sex, and significance of age group effect (reference level = Younger Age) and sex (reference level = Male)

Set work directory and load libraries
```{r}
work.dir <- "/restricted/projectnb/uh2-sebas/analysis/scCentenarians/"

library(dplyr)
library(Seurat)
library(rjags)
library(tidyverse)
library(hablar)
```

Setup data for rjags
```{r}
#load and integrate datasets
pbmc.combined <- readRDS(file = paste0(work.dir,"/Integrated_Data/pbmc.combined.rds"))

#Create table of cell type abundances across samples
cell_counts <- table(pbmc.combined@meta.data$sample.ID, pbmc.combined@meta.data$ct.consensus)

#Get Age information
sample.names <- rownames(cell_counts)

#Add age information to data

#dummy variable for the four age groups across the human lifespan 
#four age groups
pbmc.combined@meta.data$age.quartiles <- factor(pbmc.combined@meta.data$age.quartiles, levels = c("20-39", "40-59", "60-89","Extreme Longevity"))
age.group <- sapply(1:nrow(cell_counts), function(x){unique(pbmc.combined@meta.data$age.quartiles[pbmc.combined@meta.data$sample.ID %in% rownames(cell_counts)[x]])})
levels(age.group) <- c(1,2,3,4)

#dummy variable for sex
pbmc.combined@meta.data$sex <- as.factor(pbmc.combined@meta.data$sex)
pbmc.combined@meta.data$sex <-  relevel(pbmc.combined@meta.data$sex, "Male")
sex <- sapply(1:nrow(cell_counts), function(x){unique(pbmc.combined@meta.data$sex[pbmc.combined@meta.data$sample.ID %in% rownames(cell_counts)[x]])})
levels(sex) <- c(1,2)


rjags_data <- list(y = cell_counts, 
             N.sample = nrow(cell_counts),
             N.ct = ncol(cell_counts), 
             N.total = apply(cell_counts, 1, sum),
             N.age = length(levels(age.group)),
             age.group = age.group, #age group for each sample
             sex = sex # sex for each sample
)
saveRDS(rjags_data, paste0(work.dir,"/Composition_Multinomial/RJAGS_Composition_Analysis/cell_counts.rds"))

rjags_data
```

Run multinomial logistic regression model
```{r}
set.seed(562)
#load model
source(paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/model_multinom.R"))
#Run model with 500 burnin iterations and 1000 total iterations
jags <- jags.model(textConnection(multinom.model),data=rjags_data, n.adapt=500, inits = list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 3))
#monitor age group effect b1 and predicted probabilities p
coda.model <- coda.samples(jags, c("b1","b2","p"), n.iter=1000)
#coda.model <- coda.samples(jags, c("b1"), n.iter=5000)

saveRDS(coda.model, paste0(work.dir, "Composition_Multinomial/RJAGS_Composition_Analysis/rjags_coda_estimates.rds"))
```

Age group effects

- comparing Middle, Older, and EL age groups to Younger age
```{r}
coda.summary <- summary(coda.model)
#extract age effects for each cell type
#mean estimate for b1[age.group, cell type]
#Middle v. Younger
b1.pos <- grep("b1", names(coda.summary$statistics[,"Mean"]))
b1.names <- names(coda.summary$statistics[b1.pos,"Mean"])
b1.effects <- coda.summary$statistics[,c("Mean", "SD")][b1.pos,] %>%
                as_tibble(rownames = "coef") %>%
                mutate("cell.type" = c(sapply(1:rjags_data$N.ct,function(x) rep(colnames(rjags_data$y)[x],rjags_data$N.age))), 
                       "age.group" = rep(c("Younger.Ref","Middle.v.Younger", "Older.v.Younger", "EL.v.Younger"), rjags_data$N.ct)
                       ) %>%
                filter(age.group != "Younger.Ref") %>%
                mutate("zscore" = Mean/SD,
                       "pval" = 2*(1-pnorm(abs(zscore))),
                       "fdr" = p.adjust(pval,method="BH")
                       ) %>%
                dplyr::select(coef, cell.type, age.group, Mean, SD, zscore, pval, fdr)
b1.effects

#write.csv(b1.effects, paste0(work.dir, "Composition_Multinomial/RJAGS_Composition_Analysis/b1.effects.csv"))
```
Heatmap of age coefficients
```{r}
ct_order <- c("CD4 T Naive", "CD4 T Memory", "CD4 T Cytotoxic", "CD8 T", "T gamma delta", "NK","BC Naive", "BC Memory", "Plasma", "M14", "M16", "mDC", "pDC")

b1.mat <- b1.effects %>% 
  select(cell.type, age.group, Mean) %>% 
  spread(key = age.group, value = Mean) %>% 
  relocate(EL.v.Younger, .after = Older.v.Younger) %>%
   arrange(factor(cell.type, levels = ct_order)) 

b1.mat

library(circlize)
col_fun <- colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))
row_col <- c("cadetblue2",
            "cornflowerblue",
            "blue2",
            "aquamarine3",
            "turquoise4",
            "seagreen",
            "mediumpurple2",
            "purple2",
            "darkorchid4",
            "pink2",
            "red3",
            "sienna2",
            "yellow2")
names(row_col) <- ct_order
library(ComplexHeatmap)

column_ha <- HeatmapAnnotation(comparison = c("Middle.v.Younger", "Older.v.Younger", "EL.v.Younger"), col = list(comparison = c("Middle.v.Younger"= "grey","Older.v.Younger"="grey", "EL.v.Younger"="grey")))
row_ha <- rowAnnotation(cell.type = ct_order , col = list(cell.type = row_col))

b1.heatmap <- ComplexHeatmap::Heatmap(b1.mat[2:4], col = col_fun, name = "age coefficient", cluster_rows = F, cluster_columns = F, top_annotation = column_ha, left_annotation = row_ha)

b1.heatmap
```

Sex effects

- comparing Females compared to Males
```{r}
#extract sex effects for each cell type
#mean estimate for b2[sex, cell type]
b2.pos <- grep("b2", names(coda.summary$statistics[,"Mean"]))
b2.names <- names(coda.summary$statistics[b2.pos,"Mean"])
b2.effects <- coda.summary$statistics[,c("Mean", "SD")][b2.pos,] %>%
                as_tibble(rownames = "coef") %>%
                mutate("cell.type" = c(sapply(1:rjags_data$N.ct,function(x) rep(colnames(rjags_data$y)[x],2))), 
                       "sex" = rep(c("Male","Female.v.Male"), rjags_data$N.ct)
                       ) %>%
                filter(Mean != 0 & SD != 0) %>%
                mutate("zscore" = Mean/SD,
                       "pval" = 2*(1-pnorm(abs(zscore))),
                       "fdr" = p.adjust(pval,method="BH")
                       ) %>%
                dplyr::select(coef, cell.type, sex, Mean, SD, zscore, pval, fdr)
b2.effects

#write.csv(b2.effects, paste0(work.dir, "Composition_Multinomial/RJAGS_Composition_Analysis/b2.effects.csv"))
```

```{r}
b2.mat <- b2.effects %>% 
  select(cell.type, sex, Mean) %>% 
  spread(key = sex, value = Mean) %>% 
  #relocate(EL.v.Younger, .after = Older.v.Younger) %>%
   arrange(factor(cell.type, levels = ct_order)) 

b2.mat

library(circlize)
col_fun <- colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))
row_col <- c("cadetblue2",
            "cornflowerblue",
            "blue2",
            "aquamarine3",
            "turquoise4",
            "seagreen",
            "mediumpurple2",
            "purple2",
            "darkorchid4",
            "pink2",
            "red3",
            "sienna2",
            "yellow2")
names(row_col) <- ct_order
library(ComplexHeatmap)

column_ha <- HeatmapAnnotation(comparison = c("Female.v.Male"), col = list(comparison = c("Female.v.Male"= "grey")))
row_ha <- rowAnnotation(cell.type = ct_order , col = list(cell.type = row_col))

b2.heatmap <- ComplexHeatmap::Heatmap(b2.mat[2], col = col_fun, name = "sex coefficient", cluster_rows = F, cluster_columns = F, top_annotation = column_ha, left_annotation = row_ha)

b2.heatmap

```

# Extract cell type specific predicted probabilities and credible intervals

## Predicted probabilities
```{r}
#extract probabilities and create matrix with rows as samples and columns as cell types
#Extract mean and SE
predicted.estimate <- coda.summary$statistics[,c("Mean", "SD")][-c(b1.pos, b2.pos),] %>%
                as_tibble(rownames = "coef") %>%
                mutate("sample.ID" = rep(rownames(rjags_data$y), rjags_data$N.ct),
                  "cell.type" = c(sapply(1:rjags_data$N.ct,function(x) rep(colnames(rjags_data$y)[x],rjags_data$N.sample))), 
                  "age.group" = rep(rjags_data$age.group, rjags_data$N.ct),
                  "sex" = rep(rjags_data$sex, rjags_data$N.ct)
                       )
#Extract 95% credible interval
predicted.credible <- coda.summary$quantiles[,c("2.5%", "97.5%")][-c(b1.pos, b2.pos),] %>%
                as_tibble(rownames = "coef") %>%
                mutate("sample.ID" = rep(rownames(rjags_data$y), rjags_data$N.ct),
                  "cell.type" = c(sapply(1:rjags_data$N.ct,function(x) rep(colnames(rjags_data$y)[x],rjags_data$N.sample))), 
                       "age.group" = rep(rjags_data$age.group, rjags_data$N.ct),
                  "sex" = rep(rjags_data$sex, rjags_data$N.ct)
                  ) %>%
                rename(CI_2.5 = "2.5%", CI_97.5 = "97.5%")

#merge dataframes together using join
predicted.prob <- inner_join(predicted.estimate, predicted.credible, by = c("coef", "cell.type", "age.group", "sample.ID", "sex")) %>%
            hablar::convert(fct(age.group, cell.type, sex))
            
            
            
predicted.prob
```

Extract age group and sex levels
```{r}
levels(predicted.prob$age.group) <- c("Younger", "Middle", "Older", "EL")
levels(predicted.prob$sex) <- c("Male", "Female")

predicted.prob$cell.type <- factor(predicted.prob$cell.type, levels = c("CD4 T Naive", "CD4 T Memory", "CD4 T Cytotoxic", "CD8 T", "T gamma delta", "NK", "BC Naive", "BC Memory", "Plasma", "M14", "M16", "mDC", "pDC"))

levels(predicted.prob$cell.type) <- c("CD4 T Naive", "CD4 T Memory", "CD4 T Cytotoxic", "CD8 T Cytotoxic", "T Gamma Delta", "Natural Killer","B Naive", "B Memory", "Plasma", "CD14 Mono", "CD16 Mono", "Myeloid DC", "Plasmacytoid DC")

predicted.prob

#write.csv(predicted.prob, paste0(work.dir, "Composition_Multinomial/RJAGS_Composition_Analysis/predicted.probabilities.csv"))
```

# Relative proportion of cell types per age group to compare the probability estimates

```{r}
pbmc.combined <- readRDS(paste0(work.dir, "/Integrated_Data/pbmc.combined.rds"))
pbmc.combined$ct.consensus <- droplevels(pbmc.combined$ct.consensus)

Idents(pbmc.combined) <- "sex"
pbmc.male <- subset(pbmc.combined, idents = "Male")
pbmc.female <- subset(pbmc.combined, idents = "Female")

relative.prop.male <- prop.table(table(cell.type = pbmc.male@meta.data$ct.consensus, age.group = pbmc.male@meta.data$age.quartile), margin = 2) %>%
                  as_tibble() %>%
                  rename(proportion = n) %>%
                  hablar::convert(fct(cell.type, age.group)) %>%
                  mutate(sex = "Male")
levels(relative.prop.male$age.group) <- c("Younger", "Middle", "Older", "EL")


pbmc.female <- subset(pbmc.combined, idents = "Female")

relative.prop.female <- prop.table(table(cell.type = pbmc.female@meta.data$ct.consensus, age.group = pbmc.female@meta.data$age.quartile), margin = 2) %>%
                  as_tibble() %>%
                  rename(proportion = n) %>%
                  hablar::convert(fct(cell.type, age.group)) %>%
                  mutate(sex = "Female")

levels(relative.prop.female$age.group) <- c("Younger", "Middle", "Older", "EL")

relative.prop <- bind_rows(relative.prop.male, relative.prop.female) %>%
                  hablar::convert(fct(sex))
relative.prop$cell.type <- factor(relative.prop$cell.type, levels = c("CD4 T Naive", "CD4 T Memory", "CD4 T Cytotoxic", "CD8 T", "T gamma delta", "NK", "BC Naive", "BC Memory", "Plasma", "M14", "M16", "mDC", "pDC"))

  
relative.prop
```

# Plot cell type probabilities and relative proportions across age groups

## All cell types
### Estimated Proportions

```{r}

colors <- c("cadetblue2",
            "cornflowerblue",
            "blue2",
            "aquamarine3",
            "turquoise4",
            "seagreen",
            "mediumpurple2",
            "purple2",
            "darkorchid4",
            "pink2",
            "red3",
            "sienna2",
            "yellow2")

predicted.avg <- predicted.prob %>% group_by(cell.type, age.group, sex) %>% summarize(avg_proportion = mean(Mean)) %>% mutate(avg_percent = round(avg_proportion * 100, 2))

plot.barplot <- ggplot(data=predicted.avg, aes(x = age.group, y = avg_proportion, fill = cell.type, label = avg_percent)) +
        geom_bar(color = "black", stat="identity", position = "stack") + facet_wrap(~sex) +
  geom_text(aes(label=paste0(sprintf("%1.1f", avg_percent),"%")), position=position_stack(vjust=0.5), colour="black", size = 8) +
  scale_fill_manual(values = colors ) +
  theme_classic(base_size = 40) + 
  theme(axis.line=element_line()) + 
  xlab("age group") +
  ylab("estimated proportions") 
#+
 # ggtitle("Multinomial Logistic Regression Results") + theme(legend.position = "none")
#+
 # ggstyle(scale = 1)

plot.barplot

# ggplot2::ggsave(
#     filename="Figure_02C_predicted.barplot.pdf",
#     plot=plot.barplot,
#     path=paste0(work.dir,"/Figures"),
#     scale=1,
#     width = 20,
#     height=10,
#     units="in",
#     dpi=300
# )

```

### Relative Proportions

```{r}

colors <- c("cadetblue2",
            "cornflowerblue",
            "blue2",
            "aquamarine3",
            "turquoise4",
            "seagreen",
            "mediumpurple2",
            "purple2",
            "darkorchid4",
            "pink2",
            "red3",
            "sienna2",
            "yellow2")

relative.prop <- relative.prop %>% mutate(avg_percent = round(proportion * 100, 2))

plot.relative <- ggplot(data=relative.prop, aes(x = age.group, y = avg_percent, fill = cell.type, label = avg_percent)) +
        geom_bar(color = "black", stat="identity", position = "stack") + facet_wrap(~sex) +
  geom_text(aes(label=paste0(sprintf("%1.1f", avg_percent),"%")), position=position_stack(vjust=0.5), colour="black", size = 8) +
  scale_fill_manual(values = colors ) +
  theme_classic(base_size = 40) + 
  theme(axis.line=element_line()) + 
  xlab("age group") +
  ylab("proportions") +
  ggtitle("Relative Proportions (ignoring constraints)") 

#+
  #ggstyle(scale = 1)

plot.relative

# ggplot2::ggsave(
#     filename="Figure_02C_predicted.barplot.pdf",
#     plot=plot.barplot,
#     path=paste0(work.dir,"/Figures"),
#     scale=1,
#     width = 20,
#     height=10,
#     units="in",
#     dpi=300
# )

```


## CD4 T Naive
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "CD4 T Naive")) +
        geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +  
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)  # use a white background 

ggplot2::ggsave(
    filename="CD4_T_Naive.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "CD4 T Naive"), aes(age.group, proportion, color = age.group, shape = sex)) +
        geom_point(size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## CD4 T Memory
```{r, fig.width=20, fig.height=10}
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "CD4 T Memory")) +
        geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)  # use a white background 

ggplot2::ggsave(
    filename="CD4_T_Memory.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "CD4 T Memory"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## CD4 T Cytotoxic
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "CD4 T Cytotoxic")) +
         geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)   # use a white background 

ggplot2::ggsave(
    filename="CD4_T_Cytotoxic.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "CD4 T Cytotoxic"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## CD8 T Cytotoxic
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "CD8 T")) +
        geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)   # use a white background 

ggplot2::ggsave(
    filename="CD8_T_Cytotoxic.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "CD8 T"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## B Naive
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "BC Naive")) +
         geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)  # use a white background 

ggplot2::ggsave(
    filename="B_Naive.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "BC Naive"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## B Memory
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "BC Memory")) +
         geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)   # use a white background 

ggplot2::ggsave(
    filename="B_Memory.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "BC Memory"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## Plasma cells
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "Plasma")) +
         geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)   # use a white background 

ggplot2::ggsave(
    filename="Plasma.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "Plasma"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## Natural Killer
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "NK")) +
         geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)   # use a white background 

ggplot2::ggsave(
    filename="Natural_Killer.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "NK"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## T gamma delta
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "T gamma delta")) +
          geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)    # use a white background 

ggplot2::ggsave(
    filename="T_Gamma_Delta.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "T gamma delta"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## CD14 Monocyte
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "M14")) +
         geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)  # use a white background 

ggplot2::ggsave(
    filename="CD14_Mono.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "M14"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## CD16 Monocyte
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "M16")) +
         geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)   # use a white background 

ggplot2::ggsave(
    filename="CD16_Mono.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "M16"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## Myeloid DC
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "mDC")) +
         geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)  

ggplot2::ggsave(
    filename="Myeloid_DC.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "mDC"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```

## Plasmacytoid DC
```{r, fig.width=20, fig.height=10}
library(ggplot2)
plot.multinom <- ggplot(data=predicted.prob %>% filter(cell.type == "pDC")) +
         geom_pointrange(aes(x=age.group, y=Mean, ymin=CI_2.5, ymax=CI_97.5, group=age.group, color=age.group, shape = sex), size = 1) +
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, scale = 'free') +
        xlab("age group") + ylab("composition probability (95% CI)") +
        theme_bw(base_size = 30)   # use a white background 

ggplot2::ggsave(
    filename="Plasmacytoid_DC.pdf",
    plot=plot.multinom,
    path=paste0(work.dir,"Composition_Multinomial/RJAGS_Composition_Analysis/Figures/"),
    scale=1,
    width = 10,
    height=5,
    units="in",
    dpi=300
)

plot.relative <- ggplot(data=relative.prop %>% filter(cell.type == "pDC"), aes(age.group, proportion)) +
        geom_point(aes(color = age.group), size = 4)+
  scale_color_manual(values = c("blue2", "green","red", "maroon")) +
        facet_wrap(~cell.type, ncol = 4, scale = 'free') +  
        xlab("age group") + ylab("relative proportions") +
        theme_bw(base_size = 30)  # use a white background 


plot.multinom + plot.relative + patchwork::plot_layout(nrow = 1)
```


