---
title: "Seurat clustering and analysis of scRNA-seq of PBMCs from 45 healthy individuals"
author: "Tanya Karagiannis"
date: "01/16/20"
output: html_document
---

Seurat 3
R/3.6.0


Purpose: 
Reanalyze the scRNA-seq data from natgen2018 paper


Load libraries and processed data
```{r}
suppressMessages(
c(library(readr),
library(dplyr),
library(Seurat),
library(reshape2),
library(Matrix),
library(useful),
library(cowplot),
library(umap))
)
Metadata <- read.delim("/restricted/projectnb/montilab-p/CBMrepositoryData/scRNAseq/longevity/natgen2018/raw/Metadata.txt")
PBMC_45 <- readRDS(file = "/restricted/projectnb/montilab-p/CBMrepositoryData/scRNAseq/longevity/natgen2018/processed/processed_natgen2018.rds")
```

Cluster and visualize original cell types using tSNE.
Selected first 16 PCs based on significance from spree plot (parameter from original analysis)
```{r }

PBMC_45 <- RunPCA(PBMC_45, features = VariableFeatures(PBMC_45))

#PBMC_45 <- PBMC_45 %>% 
#            FindNeighbors(dims = 1:16) %>%
#            FindClusters(PBMC_45, resolution = 0.5)

PBMC_45 <- RunTSNE(PBMC_45, reduction = "pca", dims = 1:16)
#PBMC_45 <- RunUMAP(PBMC_45, reduction = "pca", dims = 1:16)

saveRDS(PBMC_45, file = "/restricted/projectnb/montilab-p/projects/longevity/natgen2018/data/natgen2018.rds")

# Projecting singlet identities on TSNE visualization
p1 <- DimPlot(PBMC_45, group.by = "updated.cell.types",reduction = "tsne")
p2 <- DimPlot(PBMC_45, group.by = "sample.ID", reduction = "tsne")
plot_grid(p1, p2)
```

Average expression of gene markers based on original cell types identified
```{r}
Idents(PBMC_45) <- "orig.cell.types"
ct <- rev(c("CD4+ T", "CD8+ T","CD56(bright) NK","CD56(dim) NK","B","cMonocyte","ncMonocyte","mDC","Megakaryocyte","Plasma","pDC"))

Idents(PBMC_45) <- factor(Idents(PBMC_45), levels = ct )

markers.to.plot <- c("CD3D", "CREM", "HSPH1", "SELL", "GIMAP5", "CACYBP", "GZMB","KLRC1","GNLY", "NKG7", "CCL5", 
                     "CD8A", "MS4A1", "CD79A", "MIR155HG", "NME1", "FCGR3A", "VMO1", "CCL2", "S100A9", "HLA.DQA1", 
                     "GPR183", "PPBP", "GNG11", "HBA2", "HBB","TSPAN13", "IL3RA", "IGJ", "MZB1")
DotPlot(PBMC_45, features = rev(markers.to.plot), dot.scale = 8) + RotatedAxis()

```

Average expression of gene markers based on updated cell type identification
-based on gsva results
```{r}
Idents(PBMC_45) <- "updated.cell.types"
ct <- rev(c("CD4+ T", "CD8+ T","CD56(bright) NK","CD56(dim) NK","B","cMonocyte","ncMonocyte","mDC","Megakaryocyte","Plasma","pDC"))

Idents(PBMC_45) <- factor(Idents(PBMC_45), levels = ct )

markers.to.plot <- c("CD3D", "CREM", "HSPH1", "SELL", "GIMAP5", "CACYBP", "GZMB","KLRC1","GNLY", "NKG7", "CCL5", 
                     "CD8A", "MS4A1", "CD79A", "MIR155HG", "NME1", "FCGR3A", "VMO1", "CCL2", "S100A9", "HLA.DQA1", 
                     "GPR183", "PPBP", "GNG11", "HBA2", "HBB","TSPAN13", "IL3RA", "IGJ", "MZB1")
DotPlot(PBMC_45, features = rev(markers.to.plot), dot.scale = 8) + RotatedAxis()

```

Cell type gene markers
```{r}
FeaturePlot(PBMC_45, features = c("CD3D","CD8A","GZMB","KLRC1","CD14","FCGR3A","MS4A1","CD79A","ITGAX","PPBP"))

```

Combine age groups in order to have similar number of samples per age group 
```{r}
Idents(PBMC_45) <- "Age"

Idents(PBMC_45, cells = WhichCells(PBMC_45, idents = c("20-29"))) <- "20-39"
Idents(PBMC_45, cells = WhichCells(PBMC_45, idents = c("30-39"))) <- "20-39"
Idents(PBMC_45, cells = WhichCells(PBMC_45, idents = c("60-69"))) <- "60-79"
Idents(PBMC_45, cells = WhichCells(PBMC_45, idents = c("70-79"))) <- "60-79"
PBMC_45$comb.age <- Idents(PBMC_45)

saveRDS(PBMC_45, file = "/restricted/projectnb/montilab-p/projects/longevity/natgen2018/data/natgen2018.rds")
```

Number of samples per age group
```{r}
table_samples <- table(Metadata$Sample, Metadata$Age.by.decade..1.1.2017.)
table_samples <- colSums(table_samples)
table_samples <- as.data.frame(table_samples)

table_samples <- tibble::rownames_to_column(table_samples, "age")
colnames(table_samples) <- c("age", "num.samples")
table_samples

ggplot(table_samples, aes(age, num.samples)) + geom_bar(stat = "identity", aes(fill = age),  color = "black") +
  scale_fill_manual(values=RColorBrewer::brewer.pal(n = 8, name = "PRGn")[c(1:3,6:8)]) + 
  geom_text(aes(label=num.samples), vjust=1.6, color="black",position = position_dodge(0.9), size=10) +
  #theme(text = element_text(size = 20)) +
  theme_classic(base_size = 20)


```


Number of cells per sample
```{r}
table_cells <- table(PBMC_45$sample.ID, PBMC_45$orig.ident)
table_cells <- as.data.frame(table_cells)

colnames(table_cells) <- c("sample.ID", "project","num.cells")

#add back age
Age <- unlist(sapply(1:nrow(table_cells), function(x){Metadata$Age.by.decade..1.1.2017.[Metadata$Sample %in% table_cells$sample.ID[x]]}))
table_cells$age <- Age

#ggplot number of cells for all samples
ggplot(table_cells, aes(age, num.cells)) + geom_bar(stat = "identity", aes(group = sample.ID, fill = age), color = "black", position = "dodge") +
  scale_fill_manual(values=RColorBrewer::brewer.pal(n = 8, name = "PRGn")[c(1:3,6:8)]) + 
  theme_classic(base_size = 20) +
  ylim(0, 1500)



```

Proportion of cell types for each sample grouped by age
```{r}

#cell type proportion for each sample
table_age <- prop.table(table(PBMC_45@meta.data$updated.cell.types, PBMC_45@meta.data$sample.ID), margin = 2)

#reshape data
table_age <- melt(table_age)
colnames(table_age) <- c("cell.types","sample.ID","prop")

#Add back age
#Set based on combined age groups
Age <- unlist(sapply(1:nrow(table_age), function(x){Metadata$Age.by.decade..1.1.2017.[Metadata$Sample %in% table_age$sample.ID[x]]}))

table_age <- as.data.frame(table_age)

table_age$age <- Age
table_age$age <- as.character(table_age$age)
table_age$age[which(table_age$age == "20-29")] <- "20-39"
table_age$age[which(table_age$age == "30-39")] <- "20-39"
table_age$age[which(table_age$age == "60-69")] <- "60-79"
table_age$age[which(table_age$age == "70-79")] <- "60-79"

#ggplot cell type proportions for each sample based on age group
p3 <- ggplot(table_age, aes(age, prop)) + geom_boxplot(aes(fill = age )) + geom_point() + facet_wrap(~cell.types, scales = "free_y") + theme_classic(base_size = 20) + theme(axis.line=element_line()) + stat_compare_means(method = "anova")
p3 <- p3 + scale_fill_manual(values=RColorBrewer::brewer.pal(n = 8, name = "PRGn")[c(1,3,6,8)])

p3

```


```{r}
sessionInfo()

```
