---
title: "Differential Analysis of cell type specific EL and Aging signatures"
author: "Tanya Karagiannis, Stefano Monti, Paola Sebastiani"
output:
  html_document:
    theme: united
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapse: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
## Functions to run to make nice figures!

#ink colors from anfed/risoink

#' Raw riso ink colors
#' 
#' @keywords internal
.riso_colors <- function() {
  c(
    "black"="#000000",
    "burgundy"="#914E72",
    "blue"="#0078BF",
    "green"="#00A95C",
    "medium-blue"="#3255A4",
    "bright-red"="#F15060",
    "risofederal-blue"="#3D5588",
    "purple"="#765BA7",
    "teal"="#00838A",
    "flat-gold"="#BB8B41",
    "hunter-green"="#407060",
    "red"="#FF665E",
    "brown"="#925F52",
    "yellow"="#FFE800",
    "marine-red"="#D2515E",
    "orange"="#FF6C2F",
    "fluorescent-pink"="#FF48B0",
    "light-gray"="#88898A",
    "metallic-gold"="#AC936E",
    "crimson"="#E45D50",
    "fluorescent-orange"="#FF7477",
    "cornflower"="#62A8E5",
    "sky-blue"="#4982CF",
    "sea-blue"="#0074A2",
    "lake"="#235BA8",
    "indigo"="#484D7A",
    "midnight"="#435060",
    "mist"="#D5E4C0",
    "granite"="#A5AAA8",
    "charcoal"="#70747C",
    "smoky-teal"="#5F8289",
    "steel"="#375E77",
    "slate"="#5E695E",
    "turquoise"="#00AA93",
    "emerald"="#19975D",
    "grass"="#397E58",
    "forest"="#516E5A",
    "spruce"="#4A635D",
    "moss"="#68724D",
    "sea-foam"="#62C2B1",
    "kelly-green"="#67B346",
    "light-teal"="#009DA5",
    "ivy"="#169B62",
    "pine"="#237E74",
    "lagoon"="#2F6165",
    "violet"="#9D7AD2",
    "orchid"="#AA60BF",
    "plum"="#845991",
    "raisin"="#775D7A",
    "grape"="#6C5D80",
    "scarlet"="#F65058",
    "tomato"="#D2515E",
    "cranberry"="#D1517A",
    "maroon"="#9E4C6E",
    "raspberry-red"="#D1517A",
    "brick"="#A75154",
    "light-lime"="#E3ED55",
    "sunflower"="#FFB511",
    "melon"="#FFAE3B",
    "apricot"="#F6A04D",
    "paprika"="#EE7F4B",
    "pumpkin"="#FF6F4C",
    "bright-olive-green"="#B49F29",
    "bright-gold"="#BA8032",
    "copper"="#BD6439",
    "mahogany"="#8E595A",
    "bisque"="#F2CDCF",
    "bubble-gum"="#F984CA",
    "light-mauve"="#E6B5C9",
    "dark-mauve"="#BD8CA6",
    "wine"="#914E72",
    "gray"="#928D88",
    "white"="#FFFFFF",
    "aqua"="#5EC8E5",
    "mint"="#82D8D5",
    "fluorescent-yellow"="#FFE900",
    "fluorescent-red"="#FF4C65",
    "fluorescent-green"="#44D62C"
  )
}

#' Adjust transparency of a hex string
#'
#' @param hex A 6-character hex string (e.g. #000000)
#' @param percent Transparency level from 0-1
#' @return A hex string
#' 
#' @export
hex_transparacy <- function(hex, percent=1) {
  if (percent < 0) percent <- 0
  if (percent > 1) percent <- 1
  percent <- toupper(as.hexmode(floor(percent * 255)))
  percent <- sprintf("%02s", percent)
  hex <- paste0(hex, percent)
  return(hex)
}

#' Riso color names
#'
#' @param sorted Optionally sort values
#' @return A character vector of names
#' 
#' @export
riso_names <- function(sorted=FALSE) {
  if (sorted) {
    return(sort(names(.riso_colors())))   
  } else {
    return(names(.riso_colors()))
  }
}

#' Get a risk ink color
#'
#' @param color Color name
#' @param alpha Color transparency
#' @return A hex string
#' 
#' @export
riso_color <- function(color, alpha=1) {
  riso_colors <- .riso_colors()
  if (color %in% names(riso_colors)) {
    hex <- riso_colors[[color]]
    if (alpha >= 1) {
      return(hex)
    } else {
      return(hex_transparacy(hex, alpha))
    }
  } else {
    cat("Riso Ink Colors: \n")
    print(riso_names(sorted=TRUE))
    stop("Invalid color")
  }
}

#' Plot one or more colors
#'
#' @param colors Character vector of hex strings
#' @param title Optional plot title
#' @return A `gg` object
#' 
#' @import ggplot2
#' @import ggforce 
#' 
#' @export
show_colors <- function(colors, title="") {
  df <- data.frame(hex=colors, y=1, x=seq(length(colors)))
  fill_values <- colors
  names(fill_values) <- colors
  ggplot(df, aes(x0=x, y0=y, r=0.4, fill=hex, expand=1)) + 
    geom_circle() + 
    scale_fill_manual(values=fill_values) +
    coord_equal() + 
    theme_void() +
    labs(title=paste0("  ", title, "\n")) + 
    xlim(0, max(7, nrow(df))+1) +
    theme(legend.position="none")
}


#Plot using ggstyle to easily change specific aspects
#from montilab/best-practices
#function to modify ggplot parameters from montilab/best-practices
ggstyle <- function(font="Helvetica", scale=1) {
  fs <- function(x) x*scale # Dynamic font scaling
  ggplot2::theme(
    plot.title = ggplot2::element_text(family=font, size=fs(26), face="bold", color="#222222"),
    plot.subtitle = ggplot2::element_text(family=font, size=fs(18), margin=ggplot2::margin(0,0,5,0)),
    plot.caption = ggplot2::element_blank(),
    legend.position = "right",
    legend.text.align = 0,
    legend.background = ggplot2::element_blank(),
    legend.title = ggplot2::element_blank(),
    legend.key = ggplot2::element_blank(),
    legend.text = ggplot2::element_text(family=font, size=fs(18), color="#222222"),
    axis.title =  ggplot2::element_text(family=font, size=fs(18), color="#222222"),
    axis.text = ggplot2::element_text(family=font, size=fs(18), color="#222222"),
    axis.text.x = ggplot2::element_text(margin=ggplot2::margin(5, b=10)),
    #axis.ticks = ggplot2::element_blank(),
    axis.line = ggplot2::element_line(color="#222222"),
    panel.grid.minor = ggplot2::element_blank(),
    panel.grid.major.y = ggplot2::element_blank(),
    panel.grid.major.x = ggplot2::element_blank(),
    panel.background = ggplot2::element_blank(),
    strip.background = ggplot2::element_rect(fill="white"),
    strip.text = ggplot2::element_text(size=fs(22), hjust=0)
  )
}

```

## Load libraries and set work directory
```{r, warning = F, message=F}
library(tidyverse)
library(ggplot2)
#library(ggVennDiagram)
#library(ggvenn)
library(stringr)
library(readxl)
library(xlsx)
library(gridExtra)
library(dplyr)
library(Seurat)

work.dir <- "/restricted/projectnb/uh2-sebas/analysis/scCentenarians/"

cell.types <- c("CD4 T Naive", "CD4 T Memory","CD8 T","NK","BC Naive","BC Memory","M14","M16","mDC")

pbmc.combined <- readRDS(file = paste0(work.dir,"Integrated_Data/pbmc.combined.rds"))

```

## Differential genes of aging and EL
### Load rjags results 
```{r}
rjags.res <- lapply(1:length(cell.types), function(i){
  doc.name <- str_replace_all(cell.types[i], fixed(" "), "_")
  DE.lifespan <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  return(DE.lifespan)
})
names(rjags.res) <- cell.types
```

### Select genes with fdr <= 0.05 and min 10% change and save to csv file
```{r}
#cell type specific significant genes with log fold change, p-value, and fdr 
DE.sig <- rjags.res %>% purrr::map(function(x) as_tibble(x, rownames = "gene")) %>%
              purrr::map(function(x) select(x, gene, MY_Mean, MY_pval, MY_fdr, OY_Mean, OY_pval, OY_fdr, EL_Mean, EL_pval, EL_fdr)) %>%
              purrr::map(function(x) filter(x, MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05)) %>%
              purrr::map(function(x) filter(x, abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1))) 

#add cell type names for each dataframe
DE.sig <- lapply(1:length(DE.sig), function(x){
  res <- DE.sig[[x]]
  res$cell.type <- rep(names(rjags.res)[x], nrow(res))
  return(res)
  })

#bind dataframes together
DE.sig.df <- DE.sig %>% bind_rows %>% 
                        relocate(cell.type, .before = gene) %>%
                        rename("MY_logFC" = MY_Mean, "OY_logFC" = OY_Mean, "EL_logFC" = EL_Mean)

#save cell type specific significant signatures to csv file
readr::write_csv(DE.sig.df, paste0(work.dir, "Differential_Analysis/DE.gene.sig.csv"))
```


### Select genes with fdr <= 0.05 and min 10% change and save to csv file
```{r}
DE.genes <- lapply(1:length(cell.types), function(i){
  doc.name <- str_replace_all(cell.types[i], fixed(" "), "_")
  
  
  DE.EL <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  
  #filter significant differential genes for each analysis by fdr significance <= 0.05
  #DE.EL <-  DE.EL %>% dplyr::filter(MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05)
  #DE.EL <-  DE.EL %>% dplyr::filter(abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1))

  Unique.EL <- DE.EL %>% select(MY_Mean, MY_SD, MY_z, MY_fdr, OY_Mean, OY_SD, OY_z, OY_fdr, EL_Mean, EL_SD, EL_z, EL_fdr)
   
   if (nrow(Unique.EL) != 0){
      Unique.EL <- Unique.EL
    } else{
      Unique.EL <- c("no specific genes")
    }
  
  #Significant genes in Middle v. Younger
  MY.sig <- DE.EL %>% filter(MY_fdr <= 0.05 & abs(MY_Mean) > log2(1.1)) %>% select(MY_Mean, MY_SD, MY_z, MY_fdr)
  MY.sig.genes <- rownames(MY.sig)
  
  if (nrow(MY.sig) != 0){
      MY.sig <- MY.sig
    } else{
      MY.sig <- c("no specific genes")
    }
  
  write.xlsx(MY.sig, file = paste0(work.dir,"Differential_Analysis/MY.DEG.fdr.xlsx"), sheetName = paste0(doc.name), append = T)

  
  
  #Significant genes in Older v. Younger
  OY.sig <- DE.EL %>% filter(OY_fdr <= 0.05 & abs(OY_Mean) > log2(1.1)) %>% select(OY_Mean, OY_SD, OY_z, OY_fdr)
  OY.sig.genes <- rownames(OY.sig)
  
  if (nrow(OY.sig) != 0){
      OY.sig <- OY.sig
    } else{
      OY.sig <- c("no specific genes")
    }
  write.xlsx(OY.sig, file = paste0(work.dir,"Differential_Analysis/OY.DEG.fdr.xlsx"), sheetName = paste0(doc.name), append = T)
  
  
  #Significant genes in EL v. Younger
  EL.sig <- DE.EL %>% filter(EL_fdr <= 0.05 & abs(EL_Mean) > log2(1.1)) %>% select(EL_Mean, EL_SD, EL_z, EL_fdr)
  EL.sig.genes <- rownames(EL.sig)
  
  if (nrow(EL.sig) != 0){
      EL.sig <- EL.sig
    } else{
      EL.sig <- c("no specific genes")
    }
  write.xlsx(EL.sig, file = paste0(work.dir,"Differential_Analysis/EL.DEG.fdr.xlsx"), sheetName = paste0(doc.name), append = T)
  
  

  #table of number of genes up and down for longevity and aging
  DE.table <- c("Middle v. Younger" = length(MY.sig.genes), "Older v. Younger" = length(OY.sig.genes), "EL v. Younger" = length(EL.sig.genes))
  
  
  #write.xlsx(Unique.EL, file = paste0(work.dir,"Differential_Analysis/Lifespan.DEG.fdr.xlsx"), sheetName = paste0(doc.name), append = T)
  
  #return(DE.table)
})

DE.genes <- do.call(rbind, DE.genes)
cell.type.names <- c("CD4 T Naive", "CD4 T Memory", "CD8 T Cytotoxic", "Natural Killer", "B Naive", "B Memory", "CD14 Mono", "CD16 Mono", "Myeloid DC")
rownames(DE.genes) <- cell.type.names
#DE.genes <- t(DE.genes)


require(gridExtra)
gridExtra::grid.table(DE.genes)
```

### Boxplot of each cell type specific DE genes across age
```{r, fig.dim = c(20,20)}
DE.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.EL <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.EL$gene <- rownames(DE.EL)
  DE.EL <- DE.EL %>% filter(MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05) %>% filter(abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1)) %>% arrange(EL_fdr)
  
  return(DE.EL$gene)  
})

plot.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  pbmc.set <- Seurat::AverageExpression(pbmc.set, features = DE.sig.genes[[x]], group.by = "sample.ID", slot = "data")
  
  pbmc.set <- reshape2::melt(pbmc.set$RNA)
  colnames(pbmc.set)[1:2] <- c("gene", "sample.ID")
 
  Age <- unlist(sapply(1:nrow(pbmc.set), function(x){unique(pbmc.combined@meta.data$age.quartiles[pbmc.combined@meta.data$sample.ID %in% pbmc.set$sample.ID[x]])}))

pbmc.set$age <- Age
pbmc.set$age <- factor(pbmc.set$age, levels = c("20-39","40-59","60-89", "Extreme Longevity"))
levels(pbmc.set$age) <- c("Younger", "Middle", "Older", "EL")

plot.sig <- ggplot(pbmc.set, aes(age,value)) + geom_boxplot(aes(col = age)) + geom_point() + facet_wrap(~ gene, scale = "free") + theme_bw(base_size = 30) + ggplot2::scale_color_manual(values = c("blue2", "green","red", "maroon")) + ggtitle(doc.name) + ylab("average expression level")


return(plot.sig) 
  
})

plot.sig.genes


```

### Boxplot of each cell type specific DE genes across age: Figure 4C, 4D, 4E
```{r, fig.dim = c(20,20)}
cell.types <- c("CD4 T Naive", "CD4 T Memory", "CD8 T", "BC Naive", "BC Memory", "M16")

DE.sig.genes <- list("CD4 T Naive" = c("STK17A", "LRRN3", "LEF1", "S100A4", "SESN3"),
                     "CD4 T Memory" = c("HLA.DPA1"),
                     "CD8 T" = c("HLA.DPB1"),
                     "BC Naive" = c("NOP53","MAP1LC3B"),
                     "BC Memory" = c("COX16"),
                     "M16" = c("CTSH", "ELOB")
)
  


plot.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.gene <- DE.sig.genes[[x]]
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  pbmc.set <- Seurat::AverageExpression(pbmc.set, features = DE.gene, group.by = "sample.ID", slot = "data", assays = "RNA")
  
  pbmc.set <- tibble::as_tibble(pbmc.set$RNA, rownames = "gene") %>% mutate(gene.name = DE.gene) %>% gather(sample.ID, avg.exprs, BU10:SCN8)
  
  #pbmc.set <- reshape2::melt(pbmc.set$RNA)
  #colnames(pbmc.set)[1:2] <- c("gene", "sample.ID")
 
  Age <- unlist(sapply(1:nrow(pbmc.set), function(x){unique(pbmc.combined@meta.data$age.quartiles[pbmc.combined@meta.data$sample.ID %in% pbmc.set$sample.ID[x]])}))

pbmc.set$age <- Age
pbmc.set$age <- factor(pbmc.set$age, levels = c("20-39","40-59","60-89", "Extreme Longevity"))
levels(pbmc.set$age) <- c("Younger", "Middle", "Older", "EL")

plot.sig <- sapply(1:length(DE.gene), function(y){
  plot.DE <- ggplot(pbmc.set %>% filter(gene.name == DE.gene[y]), aes(age,avg.exprs)) + geom_boxplot(aes(col = age)) + geom_point() + facet_wrap(~ gene.name, scale = "free") + theme_bw(base_size = 20) + ggplot2::scale_color_manual(values = c("blue2", "green","red", "maroon")) + ggtitle(doc.name) + ylab("average expression level")
  
  ggplot2::ggsave(
    filename=paste0("Figure_04_", DE.gene[y], ".pdf"),
    plot=plot.DE,
    path=paste0(work.dir,"/Figures"),
    scale=1,
    width = 8,
    height=5,
    units="in",
    dpi=300
)
  
})

  
})



```

### Heatmap of cell type specific DE signatures
```{r, eval = F}
lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  pbmc.set <- Seurat::AverageExpression(pbmc.set, features = DE.sig.genes[[x]], group.by = "sample.ID", slot = "data", assays = "RNA")
  
  Age <- unlist(sapply(1:ncol(pbmc.set$RNA), function(y){unique(pbmc.combined@meta.data$age.quartiles[pbmc.combined@meta.data$sample.ID %in% colnames(pbmc.set$RNA)[y]])}))

Age <- factor(Age, levels = c("20-39","40-59","60-89", "Extreme Longevity"))
levels(Age) <- c("Younger", "Middle", "Older", "EL")


require(ComplexHeatmap)
require(circlize)

ha = HeatmapAnnotation(
    age = Age,
    col = list(age = c("Younger" = "blue2", "Middle" = "green", "Older" = "red", "EL" = "maroon")
    )
)

pbmc.set_scaled = na.omit(t(apply(pbmc.set$RNA, 1, scale)))

h1 <- ComplexHeatmap::Heatmap(pbmc.set_scaled, cluster_rows = F, cluster_columns = T, name = "scaled average expression", show_row_names = T, show_column_names = F, column_title = doc.name, row_names_gp = gpar(fontsize = 12), top_annotation = ha)

return(h1) 
  
})



```

### Average scaled expression Heatmap of all DE genes for each cell type

```{r}
DE.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.EL <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.EL$gene <- rownames(DE.EL)
  DE.EL <- DE.EL %>% filter(MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05) %>% filter(abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1)) %>% arrange(EL_fdr)
  
  return(DE.EL$gene)  
})
DE.sig.genes <- unique(do.call(c,DE.sig.genes))

plot.heatmaps <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  Idents(pbmc.set) <- "age.quartiles"
  # Younger age
  pbmc.YA <- subset(pbmc.set, idents = "20-39")
  pbmc.YA <- Seurat::AverageExpression(pbmc.YA, features = DE.sig.genes, group.by = "sample.ID", slot = "data", assays = "RNA")
  # Middle age
  pbmc.MA <- subset(pbmc.set, idents = "40-59")
  pbmc.MA <- Seurat::AverageExpression(pbmc.MA, features = DE.sig.genes, group.by = "sample.ID", slot = "data", assays = "RNA")
   # Older age
  pbmc.OA <- subset(pbmc.set, idents = "60-89")
  pbmc.OA <- Seurat::AverageExpression(pbmc.OA, features = DE.sig.genes, group.by = "sample.ID", slot = "data", assays = "RNA")
   # EL age
  pbmc.EL <- subset(pbmc.set, idents = "Extreme Longevity")
  pbmc.EL <- Seurat::AverageExpression(pbmc.EL, features = DE.sig.genes, group.by = "sample.ID", slot = "data", assays = "RNA")

require(ComplexHeatmap)
require(circlize)

pbmc.heatmap <- cbind(pbmc.YA$RNA, pbmc.MA$RNA, pbmc.OA$RNA, pbmc.EL$RNA) 
pbmc.heatmap_scaled = t(apply(pbmc.heatmap, 1, scale))

ha = HeatmapAnnotation(
    age = c(rep("Younger", dim(pbmc.YA$RNA)[2]), rep("Middle", dim(pbmc.MA$RNA)[2]), rep("Older", dim(pbmc.OA$RNA)[2]), rep("EL", dim(pbmc.EL$RNA)[2])),
    col = list(age = c("Younger" = "blue2", "Middle" = "green", "Older" = "red", "EL" = "maroon")
    )
)

h1 <- ComplexHeatmap::Heatmap(pbmc.heatmap_scaled, cluster_rows = T, cluster_columns = F, name = "scaled average expression", show_row_names = T, show_column_names = F, column_title = doc.name, row_names_gp = gpar(fontsize = 12), top_annotation = ha, na_col = "white")

return(h1) 
  
})

plot.heatmaps

plot.heatmaps[[2]] + plot.heatmaps[[3]] + plot.heatmaps[[4]] + plot.heatmaps[[5]] + plot.heatmaps[[7]] + patchwork::plot_layout(ncol = 1)

```

### Scaled expression Heatmap (single cell level) of all DE genes for each cell type

```{r}
DE.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.EL <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.EL$gene <- rownames(DE.EL)
  DE.EL <- DE.EL %>% filter(MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05) %>% filter(abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1)) %>% arrange(EL_fdr)
  
  return(DE.EL$gene)  
})
DE.sig.genes <- unique(do.call(c,DE.sig.genes))

plot.heatmaps <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  Idents(pbmc.combined) <- "ct.consensus"
  
  plot.ct <- DoHeatmap(pbmc.combined, features = DE.sig.genes, cells = WhichCells(pbmc.combined, idents = cell.types[x]), group.by = "age.quartiles", group.colors = c("blue2", "green","red", "maroon"), group.bar = TRUE)
  
  return(plot.ct)
})
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  Idents(pbmc.set) <- "age.quartiles"
  # Younger age
  pbmc.YA <- subset(pbmc.set, idents = "20-39")
  pbmc.YA.exprs <- as.matrix(GetAssayData(pbmc.YA, features = , slot = 'data', assay = 'RNA'))[DE.sig.genes,]
  # Middle age
  pbmc.MA <- subset(pbmc.set, idents = "40-59")
  pbmc.MA.exprs <- as.matrix(GetAssayData(pbmc.MA, slot = 'data', assay = 'RNA'))

   # Older age
  pbmc.OA <- subset(pbmc.set, idents = "60-89")
  pbmc.OA.exprs <- as.matrix(GetAssayData(pbmc.OA, slot = 'data', assay = 'RNA')) 
   # EL age
  pbmc.EL <- subset(pbmc.set, idents = "Extreme Longevity")
  pbmc.EL.exprs <- as.matrix(GetAssayData(pbmc.EL, slot = 'data', assay = 'RNA'))

require(ComplexHeatmap)
require(circlize)

pbmc.heatmap <- cbind.data.frame(pbmc.YA.exprs, pbmc.MA.exprs, pbmc.OA.exprs, pbmc.EL.exprs)
pbmc.heatmap_scaled = na.omit(t(apply(pbmc.heatmap, 1, scale)))

ha = HeatmapAnnotation(
    age = c(rep("Younger", dim(pbmc.YA$RNA)[2]), rep("Middle", dim(pbmc.MA$RNA)[2]), rep("Older", dim(pbmc.OA$RNA)[2]), rep("EL", dim(pbmc.EL$RNA)[2])),
    col = list(age = c("Younger" = "blue2", "Middle" = "green", "Older" = "red", "EL" = "maroon")
    )
)

h1 <- ComplexHeatmap::Heatmap(pbmc.heatmap_scaled, cluster_rows = T, cluster_columns = F, name = "scaled average expression", show_row_names = T, show_column_names = F, column_title = doc.name, row_names_gp = gpar(fontsize = 12), top_annotation = ha, na_col = "white")

return(h1) 
  
})

plot.heatmaps

plot.heatmaps[[2]] + plot.heatmaps[[3]] + plot.heatmaps[[4]] + plot.heatmaps[[5]] + plot.heatmaps[[7]] + patchwork::plot_layout(ncol = 1)

```


```{r}
#color for cell types
require(ggforce)
colors.age <- c(riso_color("midnight"),
  riso_color("turquoise"),
    riso_color("scarlet"),
    riso_color("cornflower"),
    riso_color("burgundy"),
    riso_color("indigo"),
    riso_color("brick"),
  riso_color("pumpkin", alpha = 0.9),
  riso_color("lagoon", alpha = 0.8))

names(colors.age) <- cell.types

```

### Single Heatmap of results: Figure 4B 
```{r}
cell.types <- c("CD4 T Naive", "CD4 T Memory","CD8 T","NK","BC Naive","BC Memory","M14","M16","mDC")

DE.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.EL <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.EL$gene <- rownames(DE.EL)
  DE.EL <- DE.EL %>% filter(MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05) %>% filter(abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1)) %>% arrange(EL_fdr)
  
  return(DE.EL$gene)  
})
DE.sig.genes <- unique(do.call(c,DE.sig.genes))

#Average expression of DE genes for younger age samples across cell types
#n.YA <- colSums(table(pbmc.combined@meta.data %>% filter(age.quartiles == "20-39") %>% group_by(sample.ID, ct.consensus) %>% dplyr::summarise()))

YA.DE <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  Idents(pbmc.set) <- "age.quartiles"
  pbmc.set <- subset(pbmc.set, idents = "20-39")
  pbmc.set <- Seurat::AverageExpression(pbmc.set, features = DE.sig.genes, group.by = "sample.ID", slot = "data", assays = "RNA")
  colnames(pbmc.set$RNA) <- paste0(colnames(pbmc.set$RNA),".",doc.name)
  return(list(pbmc.set$RNA, dim(pbmc.set$RNA)[2]))
})
YA.Heatmap <- do.call(cbind, lapply(1:length(YA.DE), function(x){YA.DE[[x]][[1]]}))

n.YA <- do.call(cbind, lapply(1:length(YA.DE), function(x){YA.DE[[x]][[2]]}))

n.ct.YA <- do.call(c, lapply(1:length(cell.types), function(x){
    doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
     n.ct <- rep(doc.name, n.YA[x])
     return(n.ct)
 }))


#Middle Heatmap

MA.DE <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  Idents(pbmc.set) <- "age.quartiles"
  pbmc.set <- subset(pbmc.set, idents = "40-59")
  pbmc.set <- Seurat::AverageExpression(pbmc.set, features = DE.sig.genes, group.by = "sample.ID", slot = "data", assays = "RNA")
  colnames(pbmc.set$RNA) <- paste0(colnames(pbmc.set$RNA),".",doc.name)
 return(list(pbmc.set$RNA, dim(pbmc.set$RNA)[2]))
})
MA.Heatmap <- do.call(cbind, lapply(1:length(MA.DE), function(x){MA.DE[[x]][[1]]}))

n.MA <- do.call(cbind, lapply(1:length(MA.DE), function(x){MA.DE[[x]][[2]]}))

n.ct.MA <- do.call(c, lapply(1:length(cell.types), function(x){
    doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
     n.ct <- rep(doc.name, n.MA[x])
     return(n.ct)
 }))


##Older

OA.DE <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  Idents(pbmc.set) <- "age.quartiles"
  pbmc.set <- subset(pbmc.set, idents = "60-89")
  pbmc.set <- Seurat::AverageExpression(pbmc.set, features = DE.sig.genes, group.by = "sample.ID", slot = "data", assays = "RNA")
  colnames(pbmc.set$RNA) <- paste0(colnames(pbmc.set$RNA),".",doc.name)
 return(list(pbmc.set$RNA, dim(pbmc.set$RNA)[2]))
})
OA.Heatmap <- do.call(cbind, lapply(1:length(OA.DE), function(x){OA.DE[[x]][[1]]}))

n.OA <- do.call(cbind, lapply(1:length(OA.DE), function(x){OA.DE[[x]][[2]]}))

n.ct.OA  <- do.call(c, lapply(1:length(cell.types), function(x){
    doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
     n.ct <- rep(doc.name, n.OA[x])
     return(n.ct)
 }))

## EL

EL.DE <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  Idents(pbmc.combined) <- "ct.consensus"
  pbmc.set <- subset(pbmc.combined, idents = cell.types[x])
  Idents(pbmc.set) <- "age.quartiles"
  pbmc.set <- subset(pbmc.set, idents = "Extreme Longevity")
  pbmc.set <- Seurat::AverageExpression(pbmc.set, features = DE.sig.genes, group.by = "sample.ID", slot = "data", assays = "RNA")
  colnames(pbmc.set$RNA) <- paste0(colnames(pbmc.set$RNA),".",doc.name)
 return(list(pbmc.set$RNA, dim(pbmc.set$RNA)[2]))
})
EL.Heatmap <- do.call(cbind, lapply(1:length(EL.DE), function(x){EL.DE[[x]][[1]]}))

n.EL <- do.call(cbind, lapply(1:length(EL.DE), function(x){EL.DE[[x]][[2]]}))

n.ct.EL <- do.call(c, lapply(1:length(cell.types), function(x){
    doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
     n.ct <- rep(doc.name, n.EL[x])
     return(n.ct)
 }))



### Single heatmap with everything

DE.Heatmap <- cbind(YA.Heatmap, MA.Heatmap, OA.Heatmap, EL.Heatmap)
DE.Heatmap_scaled = na.omit(t(apply(DE.Heatmap, 1, scale)))

require(ComplexHeatmap)
require(circlize)

ha <- HeatmapAnnotation(
    cell.type = c(n.ct.YA, n.ct.MA, n.ct.OA, n.ct.EL), 
    age = c(rep("Younger", dim(YA.Heatmap)[2]), rep("Middle", dim(MA.Heatmap)[2]), rep("Older", dim(OA.Heatmap)[2]), rep("EL", dim(EL.Heatmap)[2])),
    col = list(age = c("Younger" = "blue2", "Middle" = "green", "Older" = "red", "EL" = "maroon"),
               cell.type = c("CD4_T_Naive" = "cadetblue2",
                              "CD4_T_Memory" = "cornflowerblue",
                             "CD8_T" = "aquamarine3",
                             "NK" = "seagreen",
                             "BC_Naive" = "mediumpurple2",
                             "BC_Memory" = "purple2",
                             "M14" = "pink2",
                             "M16" = "red3",
                             "mDC" = "sienna2")
            )
    )

h1 <- ComplexHeatmap::Heatmap(DE.Heatmap_scaled, cluster_rows = T, cluster_columns = F, name = "scaled average expression", show_row_names = T, show_column_names = F, row_names_gp = gpar(fontsize = 10), na_col = "white", top_annotation = ha)


h1
```





### Heatmap of union of top significant genes (based on fdr) across aging and EL
```{r, echo = F, eval = F}
 
#cell.types <- c("CD4 T Naive", "CD4 T Memory","CD8 T","NK","BC Memory","BC Naive","M14","M16","mDC","pDC")

DE.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.EL <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.EL$gene <- rownames(DE.EL)
  DE.EL <- DE.EL %>% filter(MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05) %>% filter(abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1)) %>% arrange(EL_fdr)
  
  return(DE.EL$gene)  
})
DE.sig.genes <- unique(do.call(c,DE.sig.genes))


Middle.Heatmap <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.Age <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.Age$gene <- rownames(DE.Age)
  Unique.Age <- sapply(1:length(DE.sig.genes), function(x){
    if (sum(DE.sig.genes[x] %in% DE.Age$gene) != 0){
      sig <- DE.Age$MY_Mean[DE.Age$gene %in% DE.sig.genes[x]]
      return(sig)
    }else{
      return(NA)
    }
  })
  return(Unique.Age)
})
Middle.Heatmap <- do.call(cbind, Middle.Heatmap)
rownames(Middle.Heatmap) <- DE.sig.genes
colnames(Middle.Heatmap) <- c("CD4 T Naive", "CD4 T Memory", "CD8 T Cytotoxic", "Natural Killer", "B Memory", "B Naive", "CD14 Mono", "CD16 Mono", "Myeloid DC")

Older.Heatmap <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.Age <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.Age$gene <- rownames(DE.Age)
  Unique.Age <- sapply(1:length(DE.sig.genes), function(x){
    if (sum(DE.sig.genes[x] %in% DE.Age$gene) != 0){
      sig <- DE.Age$OY_Mean[DE.Age$gene %in% DE.sig.genes[x]]
      return(sig)
    }else{
      return(NA)
    }
  })
  return(Unique.Age)
})
Older.Heatmap <- do.call(cbind, Older.Heatmap)
rownames(Older.Heatmap) <- DE.sig.genes
colnames(Older.Heatmap) <- c("CD4 T Naive", "CD4 T Memory", "CD8 T Cytotoxic", "Natural Killer", "B Memory", "B Naive", "CD14 Mono", "CD16 Mono", "Myeloid DC")


EL.Heatmap <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.EL <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.EL$gene <- rownames(DE.EL)
  Unique.EL <- sapply(1:length(DE.sig.genes), function(x){
    if (sum(DE.sig.genes[x] %in% DE.EL$gene) != 0){
      sig <- DE.EL$EL_Mean[DE.EL$gene %in% DE.sig.genes[x]]
      return(sig)
    }else{
      return(NA)
    }
  })
  return(Unique.EL)
})
EL.Heatmap <- do.call(cbind, EL.Heatmap)
rownames(EL.Heatmap) <- DE.sig.genes
colnames(EL.Heatmap) <- c("CD4 T Naive", "CD4 T Memory", "CD8 T Cytotoxic", "Natural Killer", "B Memory", "B Naive", "CD14 Mono", "CD16 Mono", "Myeloid DC")

require(circlize)
require(ComplexHeatmap)
DE.Heatmap <- cbind(Middle.Heatmap, Older.Heatmap, EL.Heatmap)
DE.Heatmap <- t(apply(2^DE.Heatmap,1,scale))
#col_fun = colorRamp2(c(-4,0,4),c("darkblue", "white", "darkred"))
#col_fun = colorRamp2(c(0,1.1,4),c("blue", "white", "red"))

ha <- HeatmapAnnotation(
    cell.type = c(rep(cell.types, 3)), 
    age = c(rep("Middle.v.Younger", 9), rep("Older.v.Younger", 9), rep("EL.v.Younger", 9)),
    col = list(age = c("Middle.v.Younger" = "green", "Older.v.Younger" = "red", "EL.v.Younger" = "maroon"))
    )

h1 <- ComplexHeatmap::Heatmap(DE.Heatmap, cluster_rows = T, cluster_columns = F, name = "scaled FC", show_row_names = T, show_column_names = T, column_title = "Cell type specific signatures across the human lifespan", row_names_gp = gpar(fontsize = 12), na_col = "white", top_annotation = ha)

h1 <- ComplexHeatmap::Heatmap(as.matrix(t(scale(t(2^Middle.Heatmap)))), cluster_rows = F, cluster_columns = F, name = "scaled FC", show_row_names = F, show_column_names = T, column_title = "Middle vs. Younger Age", row_names_gp = gpar(fontsize = 12), na_col = "white")
h2 <- ComplexHeatmap::Heatmap(as.matrix(t(scale(t(2^Older.Heatmap)))), cluster_rows = F, cluster_columns = F, name = "scaled FC", show_row_names = F,  show_column_names = T, column_title = "Older vs. Younger Age", row_names_gp = gpar(fontsize = 12), na_col = "white")
h3 <- ComplexHeatmap::Heatmap(as.matrix(t(scale(t(2^EL.Heatmap)))), cluster_rows = F, cluster_columns = F, name = "scaled FC", show_row_names = T,  show_column_names = T, column_title = "EL vs. Younger Controls", row_names_gp = gpar(fontsize = 12), na_col = "white")

#pdf(paste0(work.dir,"DE.Heatmap.pdf"),10,10)
print(h1 + h2 + h3)
#dev.off()


```

### Forest Plot for each cell type for significant genes specific to each cell type
```{r, fig.dim=c(20,10)}
Forest.sig.plot <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.EL <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.EL$gene <- rownames(DE.EL)
  DE.EL <- DE.EL %>% filter(MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05) %>% filter(abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1)) %>% arrange(EL_fdr)
  
  #Forest plot for Middle Age
  DE.EL$MY_direction[2^DE.EL$MY_Mean < 1] <- 2
  DE.EL$MY_direction[2^DE.EL$MY_Mean > 1] <- 1
  DE.EL$MY_direction <- factor(DE.EL$MY_direction)
  levels(DE.EL$MY_direction) <- c("Up", "Down")
  fp.Middle <- ggplot(data=DE.EL) +
    geom_pointrange(aes(x=gene, y=2^MY_Mean, ymin=2^MY_X2.5., ymax=2^MY_X97.5., color = MY_direction), position = position_dodge(width = 1))+
    coord_flip() +  # flip coordinates (puts labels on y axis)
    xlab("gene") + ylab("FC (95% CI)") + ggtitle("Middle v. Younger Age") +
    ylim(0,2) +
    scale_color_manual(values=c("blue", "red")) +
    geom_hline(yintercept = 1, linetype="dotted", color = "grey", size=1.5) +
    theme(legend.position = "none") +
    theme_bw(base_size = 20)  # use a white background
  fp.Middle

#Forest plot for Older Age
  DE.EL$OY_direction[2^DE.EL$OY_Mean < 1] <- 2
  DE.EL$OY_direction[2^DE.EL$OY_Mean > 1] <- 1
  DE.EL$OY_direction <- factor(DE.EL$OY_direction)
  levels(DE.EL$OY_direction) <- c("Up", "Down")
  fp.Older <- ggplot(data=DE.EL) +
    geom_pointrange(aes(x=gene, y=2^OY_Mean, ymin=2^OY_X2.5., ymax=2^OY_X97.5., color = OY_direction), position = position_dodge(width = 1))+
    coord_flip() +  # flip coordinates (puts labels on y axis)
    xlab("gene") + ylab("FC (95% CI)") + ggtitle("Older v. Younger Age") +
    ylim(0,2) +
    scale_color_manual(values=c("blue", "red")) +
    geom_hline(yintercept = 1, linetype="dotted", color = "grey", size=1.5) +
    theme(legend.position = "none") +
    theme_bw(base_size = 20)  # use a white background
  fp.Older

  #Forest plot for Older Age
  DE.EL$direction[2^DE.EL$EL_Mean < 1] <- 2
  DE.EL$direction[2^DE.EL$EL_Mean > 1] <- 1
  DE.EL$direction <- factor(DE.EL$direction)
  levels(DE.EL$direction) <- c("Up", "Down")
  fp.EL <- ggplot(data=DE.EL) +
    geom_pointrange(aes(x=gene, y=2^EL_Mean, ymin=2^EL_X2.5., ymax=2^EL_X97.5., color = direction), position = position_dodge(width = 1))+
    coord_flip() +  # flip coordinates (puts labels on y axis)
    xlab("gene") + ylab("FC (95% CI)") + ggtitle("EL v. Younger Age") +
    ylim(0,2) +
    scale_color_manual(values=c("blue", "red")) +
    geom_hline(yintercept = 1, linetype="dotted", color = "grey", size=1.5) +
    theme_bw(base_size = 20)  # use a white background
  fp.EL
  
  #Combined plots in same panel
  
  require(patchwork)
  cell.type.names <- c("CD4 T Naive", "CD4 T Memory", "CD8 T Cytotoxic", "Natural Killer", "B Memory", "B Naive", "CD14 Mono", "CD16 Mono", "Myeloid DC")

  fp.lifespan <- fp.Middle + 
                 fp.Older + 
                 fp.EL + 
                 patchwork::plot_layout(nrow = 1) +
                 plot_annotation(title =cell.type.names[x],
                                 theme = theme(plot.title = element_text(size = 30)),
                                caption = 'FDR <= 0.05 and abs(logFC) < log2(1.1)',
                                )
  
  return(fp.lifespan)  
})

lapply(1:length(cell.types), function(x){
  Forest.sig.plot[[x]]
})



```

## Check tau ratios across genes for each cell type (tau/tau.s)
```{r}
#function to trim tails based on percentile
trim_q <- function(x, lb, ub){
  x[(x > quantile(x, lb)) & (x < quantile(x, ub))]
}


#plot tau ratios
var.ratio <- lapply(1:length(cell.types), function(x){
  cell.type.names <- c("CD4 T Naive", "CD4 T Memory", "CD8 T Cytotoxic", "Natural Killer", "B Memory", "B Naive", "CD14 Mono", "CD16 Mono", "Myeloid DC")
  #read in tau ratios for each cell type
  tau.ratios <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/tau.ratios.rds"))
  #for specific cell type, trim out 1% of tails
  tau.ratios <- trim_q(tau.ratios[[cell.types[x]]] , 0.01, 0.99) 
  #plot tau ratios
  tau.ratio.plot <- ggplot(data.frame(tau.ratio = tau.ratios), aes(tau.ratio)) + geom_histogram(color="black", fill="white") + geom_vline(aes(xintercept=0), color="blue", linetype="dashed", size=1) + xlab("tau ratio") + ggtitle(cell.type.names[x])
  
  return(tau.ratio.plot)  
})

#print out tau ratios
lapply(1:length(cell.types), function(x){
  var.ratio[[x]]
})


```


```{r, eval = F, echo = F}
DE.sig.genes <- lapply(1:length(cell.types), function(x){
  doc.name <- str_replace_all(cell.types[x], fixed(" "), "_")
  DE.EL <- readRDS(paste0(work.dir,"Differential_Analysis/JAGS_Output/JAGS.Lifespan/JAGS.Summary/",doc.name,".rds"))
  DE.EL$gene <- rownames(DE.EL)
  DE.EL <- DE.EL %>% filter(MY_fdr <= 0.05 | OY_fdr <= 0.05 | EL_fdr <= 0.05) %>% filter(abs(MY_Mean) > log2(1.1) | abs(OY_Mean) > log2(1.1) | abs(EL_Mean) > log2(1.1)) %>% arrange(EL_fdr)
  
  return(DE.EL$gene)  
})
DE.sig.genes <- unique(do.call(c,DE.sig.genes))


Bulk.sig.genes <- unique(c(Middle_v_Younger_sig_genes$gene, Older_v_Younger_sig_genes$gene, EL_v_Younger_sig_genes$gene))


sc.bulk.venn <- ggvenn::ggvenn(list("single cell" = DE.sig.genes, "bulk" = Bulk.sig.genes))

gridExtra::grid.table(data.frame("Common" = DE.sig.genes[DE.sig.genes %in% Bulk.sig.genes]))


Aging.markers <- c("HSPA6", "TNFRSF1A","HMOX1","GAA","CTSH","CTSD","CTSS", 
                   "TNFRSF1B","TUBA1B","EIF4A1","CD86","MAX","PRDX6","LCK","CD28","CD69")

aging.bulk.venn <- sc.bulk.venn <- ggvenn::ggvenn(list("aging.markers" = Aging.markers, "bulk" = Bulk.sig.genes))



gridExtra::grid.table(data.frame("Common" = Aging.markers[Aging.markers %in% Bulk.sig.genes]))
                      
#aging markers are found in the bulk analysis
middle.bulk.venn <- ggvenn::ggvenn(list("aging.markers" = Aging.markers, "bulk.Middle.v.Younger" = Middle_v_Younger_sig_genes$gene))
older.bulk.venn <- ggvenn::ggvenn(list("aging.markers" = Aging.markers, "bulk.Older.v.Younger" = Older_v_Younger_sig_genes$gene))
EL.bulk.venn <- ggvenn::ggvenn(list("aging.markers" = Aging.markers, "bulk.EL.v.Younger" = EL_v_Younger_sig_genes$gene))

logFC.EL.aging <- EL_v_Younger_sig_genes[EL_v_Younger_sig_genes$gene %in%  Aging.markers[Aging.markers %in% Bulk.sig.genes], ]

aging.sc.venn <- ggvenn::ggvenn(list("aging.markers" = Aging.markers, "single cell" = DE.sig.genes))

Aging.markers[Aging.markers %in% DE.sig.genes]
```