---
title: "Bulk DE analysis of single cell data at the sample level using DESeq2"
author: "Tanya Karagiannis, Stefano Monti, Paola Sebastiani"
output:
  html_document:
    theme: united
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapse: false
---

# Load libraries and data
```{r}
suppressMessages(c(
library(dplyr),
library(reshape2),
library(Seurat),
library(Matrix),
library(ggplot2),
library(cowplot),
library(knitr),
library(SingleCellExperiment),
library(Matrix.utils),
library(DESeq2),
library(tibble)
))

work.dir <- "/restricted/projectnb/uh2-sebas/analysis/scCentenarians/"

pbmc.combined <- merge(readRDS(file = paste0(work.dir,"PublicData_Reanalysis/Data/pbmc_public.rds")), readRDS(paste0(work.dir,"NECS_Analysis/Data/pbmc.necs.rds")), project = "PBMC.3data")
```

# Filter dataset:
- keep immune cell type only

```{r}
#keep only immune cells
Idents(pbmc.combined) <- "ct.consensus"
pbmc.combined <- subset(pbmc.combined, idents = c("EC","MKI","MGK","Plasma", "CD4 T Cytotoxic", "T gamma delta","pDC"), invert = TRUE)
pbmc.combined$ct.consensus <- factor(pbmc.combined$ct.consensus, levels = c("CD4 T Naive", "CD4 T Memory", "NK", "M14", "BC Memory","BC Naive","CD8 T","M16", "mDC"))
cell.types <- levels(pbmc.combined$ct.consensus)
print(cell.types)

#filtering genes to include genes expressed in over 50% of cells of the least frequent cell type
min.count <- sum(pbmc.combined@meta.data$ct.consensus == "mDC")/2

# Output a logical vector for every gene on whether the more than zero counts per cell
nonzero <- pbmc.combined[['RNA']]@counts > 0

# Sums all TRUE values and returns TRUE if more than min.count TRUE values per gene
keep_genes <- Matrix::rowSums(nonzero) >= min.count
length(keep_genes)

#filter genes based on threshold
pbmc.combined[["RNA"]]@data <- pbmc.combined[["RNA"]]@data[keep_genes,]
dim(pbmc.combined)

```

# Set up data into a SingleCellExperiment

```{r}
# Extract raw counts and metadata to create SingleCellExperiment object
counts <- pbmc.combined@assays$RNA@counts 

metadata <- pbmc.combined@meta.data

# Create single cell experiment object
necs.sce <- SingleCellExperiment(assays = list(counts = counts), 
                           colData = metadata)

#factor sample.ID
colData(necs.sce)[,c("sample.ID")] <- factor(colData(necs.sce)[,c("sample.ID")])

```

# Generate sample level metadata

```{r}

# Named vector of sample names
sids <- levels(necs.sce$sample.ID)

## Determine the number of cells per sample
table(necs.sce$sample.ID)

## Turn named vector into a numeric vector of number of cells per sample
n_cells <- as.numeric(table(necs.sce$sample.ID))

## Create the sample level metadata by combining the reordered metadata with the number of cells corresponding to each sample.
sample_metadata <- data.frame(sids, n_cells, row.names = NULL) 

#Add age information to data
#Set based on age groups: 4 control quartiles, and centenarians
Age <- unlist(sapply(1:nrow(sample_metadata), function(x){unique(pbmc.combined@meta.data$age.quartiles[pbmc.combined@meta.data$sample.ID %in% sample_metadata$sid[x]])}))

sample_metadata$age <- Age
sample_metadata$age <- factor(sample_metadata$age, levels = c("20-39","40-59","60-89", "Extreme Longevity"))
levels(sample_metadata$age) <- c("Younger","Middle","Older","EL")

#Add sex information
Sex <- unlist(sapply(1:nrow(sample_metadata), function(x){unique(pbmc.combined@meta.data$sex[pbmc.combined@meta.data$sample.ID %in% sample_metadata$sid[x]])}))

sample_metadata$sex <- Sex
sample_metadata$sex <- factor(sample_metadata$sex, levels = c("Female","Male"))

#Add batch information
Batch <- unlist(sapply(1:nrow(sample_metadata), function(x){unique(pbmc.combined@meta.data$batch[pbmc.combined@meta.data$sample.ID %in% sample_metadata$sid[x]])}))

sample_metadata$batch <- Batch
sample_metadata$batch <- factor(sample_metadata$batch, levels = c("NATGEN","COHORT.B1","COHORT.B2", "PNAS"))

#Add ethnicity information
Ethnicity <- unlist(sapply(1:nrow(sample_metadata), function(x){unique(pbmc.combined@meta.data$batch[pbmc.combined@meta.data$sample.ID %in% sample_metadata$sid[x]])}))

sample_metadata$ethnic <- Ethnicity
sample_metadata$ethnic[sample_metadata$ethnic == "NATGEN"] <- "European"
sample_metadata$ethnic[sample_metadata$ethnic == "COHORT.B1"] <- "European"
sample_metadata$ethnic[sample_metadata$ethnic == "COHORT.B2"] <- "European"
sample_metadata$ethnic[sample_metadata$ethnic == "PNAS"] <- "Japanese"
sample_metadata$ethnic <- factor(sample_metadata$ethnic, levels = c("European", "Japanese"))
#print sample metdata table
sample_metadata
```


```{r}
# Aggregate the counts per sample_id and cluster_id

# Identify groups for aggregation of counts
sample.groups <- colData(necs.sce)[, c("sample.ID")]

# Aggregate across cluster-sample groups
pb <- Matrix.utils::aggregate.Matrix(t(counts(necs.sce)), 
                       groupings = sample.groups, fun = "sum") 

class(pb)

dim(pb)

pb[1:6, 1:6]
```

# Create DEseq object

```{r}
dds <- DESeq2::DESeqDataSetFromMatrix(data.frame(t(pb)), 
                              colData = sample_metadata, 
                              design = ~ age + sex + batch)

```

# Run DESEQ at the sample level
```{r}
dds <- DESeq(dds)
#saveRDS(dds, file = paste0(work.dir,"Differential_Analysis/Bulk_Analysis/DEseq.output.rds"))
```

# Check fit of model by looking at plot of dispersion estimates
```{r}
# Plot dispersion estimates
plotDispEsts(dds)

```

# Middle Age v. Younger Age
- fdr < 0.05
- summary of results
```{r}
# Output results of Wald test for contrast for Middle v. Younger Age
levels(sample_metadata$age)[2]
levels(sample_metadata$age)[1]

contrast <- c("age", levels(sample_metadata$age)[2], levels(sample_metadata$age)[1])

# resultsNames(dds)
res <- results(dds, 
               contrast = contrast,
               alpha = 0.05)

write.csv(res %>%
        data.frame() %>%
        tibble::rownames_to_column(var="gene") ,
          paste0(work.dir, "Differential_Analysis/Bulk_Analysis/Middle.v.Younger_all_genes.csv"),
          quote = FALSE,
          row.names = FALSE)

#summary of results
summary(res)

#number of significant genes based on fdr < 0.05
sum(res$padj < 0.05, na.rm=TRUE)
```

Table of significant genes
```{r}
# Turn the results object into a tibble for use with tidyverse functions
sig.MY <- res %>%
        data.frame() %>%
        tibble::rownames_to_column(var="gene") %>%
        dplyr::filter(padj < 0.05 & abs(log2FoldChange) > log2(1.5)) %>%
        dplyr::arrange(padj) %>%
        as_tibble()

# Check results output
sig.MY

# Write all results to file
write.csv(sig.MY,
          paste0(work.dir, "Differential_Analysis/Bulk_Analysis/Middle.v.Younger_sig_genes.csv"),
          quote = FALSE,
          row.names = FALSE)
```

# Older Age v. Younger Age
- fdr < 0.05
- summary of results
```{r}
# Output results of Wald test for contrast for Older v. Younger Age
levels(sample_metadata$age)[3]
levels(sample_metadata$age)[1]

contrast <- c("age", levels(sample_metadata$age)[3], levels(sample_metadata$age)[1])

# resultsNames(dds)
res <- results(dds, 
               contrast = contrast,
               alpha = 0.05)

write.csv(res %>%
        data.frame() %>%
        tibble::rownames_to_column(var="gene") ,
          paste0(work.dir, "Differential_Analysis/Bulk_Analysis/Older.v.Younger_all_genes.csv"),
          quote = FALSE,
          row.names = FALSE)

#summary of results
summary(res)

#number of significant genes based on fdr < 0.05
sum(res$padj < 0.05, na.rm=TRUE)
```

Table of significant genes
```{r}
# Turn the results object into a tibble for use with tidyverse functions
sig.OY <- res %>%
        data.frame() %>%
        tibble::rownames_to_column(var="gene") %>%
        dplyr::filter(padj < 0.05 & abs(log2FoldChange) > log2(1.5)) %>%
        dplyr::arrange(padj) %>%
        as_tibble()

# Check results output
sig.OY

# Write all results to file
write.csv(sig.OY,
          paste0(work.dir, "Differential_Analysis/Bulk_Analysis/Older.v.Younger_sig_genes.csv"),
          quote = FALSE,
          row.names = FALSE)

```

# EL v. Younger Age
- fdr < 0.05
- summary of results
```{r}
# Output results of Wald test for contrast for EL v. Younger Age
levels(sample_metadata$age)[4]
levels(sample_metadata$age)[1]

contrast <- c("age", levels(sample_metadata$age)[4], levels(sample_metadata$age)[1])

# resultsNames(dds)
res <- results(dds, 
               contrast = contrast,
               alpha = 0.05)

write.csv(res %>%
        data.frame() %>%
        tibble::rownames_to_column(var="gene") ,
          paste0(work.dir, "Differential_Analysis/Bulk_Analysis/EL.v.Younger_all_genes.csv"),
          quote = FALSE,
          row.names = FALSE)

#summary of results
summary(res)

#number of significant genes based on fdr < 0.05
sum(res$padj < 0.05, na.rm=TRUE)
```

Table of significant genes
```{r}
# Turn the results object into a tibble for use with tidyverse functions
sig.EL <- res %>%
        data.frame() %>%
        tibble::rownames_to_column(var="gene") %>%
        dplyr::filter(padj < 0.05 & abs(log2FoldChange) > log2(1.5)) %>%
        dplyr::arrange(padj) %>%
        as_tibble()

# Check results output
sig.EL

# Write all results to file
write.csv(sig.EL,
          paste0(work.dir, "Differential_Analysis/Bulk_Analysis/EL.v.Younger_sig_genes.csv"),
          quote = FALSE,
          row.names = FALSE)

```

# Heatmap of scaled fold change of significant genes across the human lifespan
```{r}
MY_res <- read.csv(paste0(work.dir,"Differential_Analysis/Bulk_Analysis/Middle.v.Younger_all_genes.csv"), header = T)
OY_res <- read.csv(paste0(work.dir, "Differential_Analysis/Bulk_Analysis/Older.v.Younger_all_genes.csv"), header = T)
EL_res <- read.csv(paste0(work.dir,"Differential_Analysis/Bulk_Analysis/EL.v.Younger_all_genes.csv"), header = T)


DE.sig.genes <- unique(c(sig.MY$gene, sig.OY$gene, sig.EL$gene))

MY_sig_res <- MY_res[MY_res$gene %in% DE.sig.genes,] %>% select(gene, log2FoldChange) 
OY_sig_res <- OY_res[OY_res$gene %in% DE.sig.genes,] %>% select(gene, log2FoldChange) 
EL_sig_res <- EL_res[EL_res$gene %in% DE.sig.genes,] %>% select(gene, log2FoldChange) 

bulk_sig_res <- cbind.data.frame(MY_FC = MY_sig_res[,2], OY_FC = OY_sig_res[,2], EL_FC = EL_sig_res[,2])

require(circlize)
require(ComplexHeatmap)
bulk_sig_res <- scale(t(2^bulk_sig_res))

ha <- HeatmapAnnotation(comparison = c("Middle.v.Younger", "Older.v.Younger", "EL.v.Younger"), col = list(comparison = c("Middle.v.Younger" = "lightgreen", "Older.v.Younger" = "darkgreen", "EL.v.Younger" = "orchid")
))

DE.bulk.heatmap <- ComplexHeatmap::Heatmap(as.matrix(t(bulk_sig_res)), cluster_rows = T, cluster_columns = F, name = "scaled FC", show_row_names = F, show_column_names = F, row_names_gp = gpar(fontsize = 12), top_annotation = ha)

ggplot2::ggsave(
    filename="Supplementary_Bulk.DE.Heatmap.pdf",
    plot=draw(DE.bulk.heatmap),
    path=paste0(work.dir,"/Figures"),
    scale=1,
    width = 5,
    height=5,
    units="in",
    dpi=300
)
```