---
title: "Cell Type Distribution Score of aging and longevity: Keeping Naive and Memory Subtypes separate"
author: "Tanya Karagiannis, Stefano Monti, Paola Sebastiani"
output:
  html_document:
    theme: united
    toc: yes
  html_notebook:
    toc: yes
---

# Load libraries and data
```{r}
suppressMessages(c(
library(dplyr),
library(reshape2),
library(Seurat),
library(Matrix),
library(ggplot2),
library(cowplot),
library(knitr)
))

work.dir <- "/restricted/projectnb/uh2-sebas/analysis/scCentenarians/"

pbmc.combined <- readRDS(file = paste0(work.dir,"Integrated_Data/pbmc.combined.rds"))
```

# Filter dataset:
- keep immune cell type only

# Create table of cell type proportions per sample (cell types vs samples)
```{r}
ct_prop <- prop.table(table(pbmc.combined@meta.data$ct.consensus, pbmc.combined@meta.data$sample.ID), margin = 2)

require(knitr)
require(kableExtra)
kable(ct_prop, caption = paste0("cell type proportions per sample"),"html") %>% kable_styling("striped") %>% scroll_box(width = "100%",height = "300px")
```

# Histogram of cell type proportions
```{r, eval = F}
print(paste("cell.types:", rownames(ct_prop)))
p1 <- apply(ct_prop, 1, function(x){hist(x, main = rownames(ct_prop)[x])})
```

# Check that for each sample, cell type proportions add up to 1 
```{r}
apply(ct_prop,2,sum)
```

# Change data to long format and add Metadata information
```{r}
#reshape data to long format
ct_prop <- melt(ct_prop)
colnames(ct_prop) <- c("cell.types","sample.ID","proportion")


#Add age information to data
#Set based on age groups: 4 control quartiles, and centenarians
Age <- unlist(sapply(1:nrow(ct_prop), function(x){unique(pbmc.combined@meta.data$age.quartiles[pbmc.combined@meta.data$sample.ID %in% ct_prop$sample.ID[x]])}))

ct_prop$age <- Age
ct_prop$age <- factor(ct_prop$age, levels = c("20-39","40-59","60-89", "Extreme Longevity"))
levels(ct_prop$age) <- c("Younger","Middle","Older","EL")


#age by decade
age_decade_natgen <- readr::read_delim("/restricted/projectnb/uh2-sebas/data/Transcriptomics/singleCell_aging/natgen2018/raw/Metadata.txt", delim = "\t")
age_decade_natgen <- age_decade_natgen %>% select(Sample, `Age by decade (1-1-2017)`) %>% rename(decade = `Age by decade (1-1-2017)`)

age_decade_pnas <- tibble(Sample = c("SC1", "SC2", "SC3", "SC4", "SC5", "SC6", "SC7", "CT1", "CT2", "CT3", "CT4", "CT5"), decade = c(rep("110-119", 7), "50-59", "70-79","60-69","70-79", "80-89"))

age_decade_necs <- tibble(Sample = c("SCN1", "SCN2", "SCN4", "SCN5", "SCN6", "SCN7", "SCN8", "BU6", "BU10"), decade = c(rep("100-109", 2), "110-119", rep("100-109", 4),"40-49","30-39"))

age_decade <- dplyr::bind_rows(age_decade_natgen, age_decade_pnas, age_decade_necs)

Age <- unlist(sapply(1:nrow(ct_prop), function(x){unique(age_decade$decade[age_decade$Sample %in% ct_prop$sample.ID[x]])}))

ct_prop$age.group <- Age
ct_prop$age.group <- factor(ct_prop$age.group, levels = c("20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "100-109", "110-119"))



#Add sex information
Sex <- unlist(sapply(1:nrow(ct_prop), function(x){unique(pbmc.combined@meta.data$sex[pbmc.combined@meta.data$sample.ID %in% ct_prop$sample.ID[x]])}))

ct_prop$sex <- Sex

#Add batch information
Batch <- unlist(sapply(1:nrow(ct_prop), function(x){unique(pbmc.combined@meta.data$batch[pbmc.combined@meta.data$sample.ID %in% ct_prop$sample.ID[x]])}))

ct_prop$batch <- Batch
ct_prop$batch[ct_prop$batch == "COHORT.B1"] <- "NECS.B1"
ct_prop$batch[ct_prop$batch == "COHORT.B2"] <- "NECS.B2"

#Add ethnicity information
Ethnicity <- unlist(sapply(1:nrow(ct_prop), function(x){unique(pbmc.combined@meta.data$batch[pbmc.combined@meta.data$sample.ID %in% ct_prop$sample.ID[x]])}))

Ethnicity[which(Ethnicity == "COHORT.B1")] <- "European"
Ethnicity[which(Ethnicity == "COHORT.B2")] <- "European"
Ethnicity[which(Ethnicity == "NATGEN")] <- "European"
Ethnicity[which(Ethnicity == "PNAS")] <- "Japanese"

ct_prop$ethnic <- Ethnicity

#Add offspring vs. unrelated information, and centenarians vs. supercentenarians
ct_prop$relation <- "unrelated"
ct_prop$relation[ct_prop$age == "Extreme Longevity" & ct_prop$ethnic == "Japanese"] <- "supercentenarian"
ct_prop$relation[ct_prop$age == "Extreme Longevity" & ct_prop$ethnic == "European"] <- "centenarian"
ct_prop$relation[ct_prop$sample.ID == "SCN4"] <- "supercentenarian"
ct_prop$relation[ct_prop$sample.ID == "CT2"] <- "offspring"
ct_prop$relation[ct_prop$sample.ID == "CT5"] <- "offspring"
ct_prop$relation <- factor(ct_prop$relation, levels = c("unrelated", "offspring", "centenarian", "supercentenarian"))

write.csv(ct_prop, "CellTypeProportions.Subtypes.csv")

```

```{r, echo=F}
#source('~/risoink.R')
require(ggforce)
# colors <- c(riso_color("midnight"),
#   riso_color("turquoise"),
#     riso_color("violet"),
#     riso_color("scarlet"),
#     riso_color("apricot"),
#     riso_color("cornflower"),
#     riso_color("burgundy"),
#     riso_color("emerald"),
#     riso_color("indigo"),
#     riso_color("brick"),
#   riso_color("pumpkin", alpha = 0.9),
#   riso_color("lagoon", alpha = 0.8),
#   riso_color("raspberry-red"),
#   riso_color("orchid"))
# 
# show_colors(colors, "Halcyon-esque")

colors <- c("cadetblue2",
            "cornflowerblue",
            "blue2",
            "aquamarine3",
            "turquoise4",
            "seagreen",
            "mediumpurple2",
            "purple2",
            "darkorchid4",
            "pink2",
            "red3",
            "sienna2",
            "yellow2")

```

# Plot cell type proportions for all samples grouped by age group
```{r, fig.dim = c(20, 10)}
ct_prop$cell.types <- factor(ct_prop$cell.types, levels=c("CD4 T Naive","CD4 T Memory","CD4 T Cytotoxic","CD8 T","T gamma delta", "NK","BC Naive","BC Memory","Plasma", "M14","M16","mDC","pDC"))
levels(ct_prop$cell.types) <- c("CD4 T Naive", "CD4 T Memory", "CD4 T Cytotoxic", "CD8 T Cytotoxic", "T Gamma Delta", "Natural Killer","B Naive", "B Memory", "Plasma", "CD14 Mono", "CD16 Mono", "Myeloid DC", "Plasmacytoid DC")

age.groups <- levels(ct_prop$age)

#previous colors used c(colorspace::diverge_hcl(20)[1:9], rev(colorspace::diverge_hcl(20)[16:20]))

#c(colorspace::terrain_hcl(13)[1:8],colorspace::heat_hcl(13)[1:5])
p1 <- ggplot(ct_prop %>% filter(age == age.groups[1]), aes(interaction(sample.ID, age.group), proportion)) +
  geom_bar(aes(group = cell.types, fill = cell.types ), color = "black", stat="identity", position = "stack") +
  scale_fill_manual(values = colors ) +
  theme_classic(base_size = 40) + 
  theme(axis.line=element_line()) + 
  xlab("sample") +
  ylab("cell type proportions in PBMCs") +
  facet_wrap(~ age) + 
  ggstyle(scale = 1) + 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), legend.position = "none") 
p2 <- ggplot(ct_prop %>% filter(age == age.groups[2]), aes(interaction(sample.ID, age.group), proportion)) +
  geom_bar(aes(group = cell.types, fill = cell.types ), color = "black", stat="identity", position = "stack") +
 scale_fill_manual(values = colors) +
  theme_classic(base_size = 40) + 
  theme(axis.line=element_line()) + 
  xlab("sample") +
  facet_wrap(~ age) + 
  ggstyle(scale = 1) + 
  theme(axis.title.y=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "none") 
p3 <- ggplot(ct_prop %>% filter(age == age.groups[3]), aes(interaction(sample.ID, age.group), proportion)) +
  geom_bar(aes(group = cell.types, fill = cell.types ), color = "black", stat="identity", position = "stack") +
  scale_fill_manual(values = colors) +
  theme_classic(base_size = 40) + 
  xlab("sample") +
  theme(axis.line=element_line()) + 
  facet_wrap(~ age) + ggstyle(scale = 1) + 
  theme(axis.title.y=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "none") 
p4 <- ggplot(ct_prop %>% filter(age == age.groups[4]), aes(interaction(sample.ID, age.group), proportion)) +
  geom_bar(aes(group = cell.types, fill = cell.types ), color = "black", stat="identity", position = "stack") +
  scale_fill_manual(values = colors) +
  theme_classic(base_size = 40) + 
  theme(axis.line=element_line()) + 
  xlab("sample") +
  facet_wrap(~ age) + 
  ggstyle(scale = 1) + 
  theme(axis.title.y=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) 

require(patchwork)
norm.plot <- p1 + p2 + p3 + p4 + patchwork::plot_layout(nrow = 1, byrow = FALSE)

ggplot2::ggsave(
    filename="Figure_02B_subtype.prop.pdf",
    plot=norm.plot,
    path=paste0(work.dir,"/Figures"),
    scale=1,
    width = 20,
    height=8,
    units="in",
    dpi=300
)
```

# Calculate cell type distribution score for each sample
```{r, fig.dim = c(20, 10)}
#change data back to wide format: samples vs cell types
dat <- reshape(ct_prop, direction = "wide", idvar=c("sample.ID","age", "sex", "batch", "ethnic", "relation", "age.group"), timevar="cell.types")
rownames(dat) <- dat$sample.ID
dat <- dat[-c(1:7)]

#check that each samples cell type proportions add up to 1
apply(dat, 1, sum)

#apply entropy calculation to each sample to calculate cell type diversity
div.res <- apply(dat, 1, function(x){(-sum(x*log(x), na.rm = T)/log(ncol(dat))-1)})

#get wide format back
dat <- reshape(ct_prop, direction = "wide", idvar=c("sample.ID","age", "sex", "batch", "ethnic", "relation", "age.group"), timevar="cell.types")

#create matrix with entropy measures and age group information
div.res <- data.frame(entropy = div.res, dat[2:7])
write.csv(div.res, file = "CellType.Diversity.Subtypes.csv")
```

# Statistical tests
- Anova and T-test of cell type diversity between controls and extreme longevity subjects
- Multiple linear regression to look at trend of cell type diversity based on age
```{r, fig.dim = c(20, 10)}
#anova
lm.model <- lm(entropy ~ age, data = div.res)
anova(lm.model)

#pairwsie t-tests for multiple comparisons
pairwise.t.test(div.res$entropy, div.res$age, p.adjust.method = "BH")

```

```{r, fig.dim = c(20, 10)}
#Box plots of cell type diversity of samples grouped by age group and colored by age group

plot.ctds <- ggplot(div.res, aes(age, entropy)) + geom_boxplot(aes(fill = age)) + 
  theme_classic(base_size = 20) + 
  theme(axis.line=element_line()) + 
  ylim(-1, 0) +
  ylab("cell type diversity stastic") +
  scale_fill_manual(values = c("blue2", "green","red", "maroon")) +
  ggstyle() + 
  theme(axis.title.x=element_blank(),
        legend.position = "none") +
  ggsignif::geom_signif(
    y_position = c(0.0), xmin = c(0.5), xmax = c(4.5),
    annotation = c("NS"), tip_length = 0
  )

plot.ctds

ggplot2::ggsave(
    filename="Figure_Supp02_ctds.subtypes.pdf",
    plot=plot.ctds,
    path=paste0(work.dir,"/Figures"),
    scale=1,
    width = 15,
    height=8,
    units="in",
    dpi=300
)



#Box plots of cell type diversity of samples grouped by age group and colored by batch
#make sure differences are not affected by batch effects
require(ggplot2)
p5 <- ggplot(div.res, aes(age, entropy)) + geom_point(aes(fill = batch),colour = "black",pch = 21,size=3, position = position_dodge(0.3)) + theme_classic(base_size = 30) + theme(axis.line=element_line()) + ylim(0.8, 2.5) + ggstyle()
p5
```


