---
title: "Single cell analysis of CITE-seq data of centenarians from NECS Study: Final cell types identified and silhoutte scores"
author: "Tanya Karagiannis, Stefano Monti, Paola Sebastiani"
output:
  html_document:
    theme: united
    toc: yes
  html_notebook:
    toc: yes
---

```{r}
knitr::opts_chunk$set(fig.width=12, fig.height=8) 

```

```{r}
#Functions and genesets
source("/rprojectnb2/montilab-p/personal/tanya/functions/vlnplot.functions.R")
immune.genesets <- readRDS("/rprojectnb2/montilab-p/personal/tanya/functions/genesets/immune.genesets.rds")

```

```{r data, include=FALSE}
library(dplyr)
library(Seurat)
library(ggplot2)
library(reshape2)

work.dir <- "/restricted/projectnb/montilab-p/projects/longevity/scCentenarians/scCentenarians/NECS_Analysis/"

pbmc.necs <- readRDS(paste0(work.dir,"Data/pbmc.necs.rds"))
#factor cell types
cell.types <- c("CD4 T Naive", "CD4 T Memory","CD4 T Cytotoxic","CD8 T","NK","BC Naive","BC Memory","M14","M16","mDC","pDC", "MGK")

pbmc.necs$ct.consensus <- factor(pbmc.necs$ct.consensus, levels = cell.types)
```

Average silhoutte score based on batch for each cell type
- Before correction
- After Harmony correction in PCA space
```{r,eval = F}
batch_sil <- function(pca.data, batch, nPCs = 20){
  require(cluster)
    # in scone, they use svd to compute principal components.
    # For now, we'll keep the PCA object created in
    # prcomp to be consistent
    # proj <- svd(scale(t(expr),center = TRUE,scale = TRUE),
    #             nu = 3, nv =0)$u

    dd <- as.matrix(dist(pca.data[, seq_len(nPCs)]))
    summary(silhouette(as.numeric(batch), dd))$avg.width
}

cell.types <- levels(pbmc.necs$ct.consensus)
#pca.data <- pbmc@reductions$pca@cell.embeddings

pca.sil.score <- lapply(1:length(cell.types), function(x){
  ct.order <- which(pbmc.necs@meta.data$ct.consensus == cell.types[x])
  ct.cells <- rownames(pbmc.necs@meta.data)[ct.order]
  pca.data <- pbmc.necs@reductions$pca@cell.embeddings[ct.cells,]
  batch <- as.factor(pbmc.necs@meta.data$batch[ct.order])

  pca.sil <- batch_sil(pca.data, batch, nPCs = 20)
  return(pca.sil)
})
pca.sil.score <- unlist(pca.sil.score)
names(pca.sil.score) <- cell.types


harmony.sil.score <- lapply(1:length(cell.types), function(x){
  ct.order <- which(pbmc.necs@meta.data$ct.consensus == cell.types[x])
  ct.cells <- rownames(pbmc.necs@meta.data)[ct.order]
  pca.data <- pbmc.necs@reductions$harmony@cell.embeddings[ct.cells,]
  batch <- as.factor(pbmc.necs@meta.data$batch[ct.order])

  pca.sil <- batch_sil(pca.data, batch, nPCs = 20)
  return(pca.sil)
})
harmony.sil.score <- unlist(harmony.sil.score)
names(harmony.sil.score) <- cell.types

sil.score <- data.frame(pre = pca.sil.score, post = harmony.sil.score)
write.csv(sil.score, file = paste0(Data/"sil.score.NECS.csv"))

sil.score <- reshape2::melt(sil.score)
ggplot2::ggplot(sil.score, aes(
  x = variable, y = value,
  fill = variable), color = "black") +
  ggplot2::geom_boxplot() +
  ylab("silhoutte score") + xlab("batch")+
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#significance between silhoutte scores pre and post batch correction
t.test(value ~ variable, data = sil.score)

```

Visualizing final cell types and samples for NECS single cell dataset
- consensus cell types
- centenarians vs. control age groups
```{r umap}
#cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
#age.group <- c(cbPalette[6], cbPalette[8])


#Visualizing UMAP of consensus cell types
DimPlot(pbmc.necs, pt.size = 1,group.by = "ct.consensus")

#visualizing UMAP of samples
#sample.ID colors
samp.col <- colorRamps::primary.colors(9, steps = 2)

pbmc.necs$age.group <- factor(pbmc.necs$age.group, levels = c("CT", "EL"))
DimPlot(pbmc.necs, group.by = "sample.ID")

#number of cell per sample
dat <- melt(apply(table(pbmc.necs$sample.ID, pbmc.necs$orig.ident),1,sum))
dat$sample.ID <- rownames(dat)
ggplot(dat, aes(sample.ID, value, fill = sample.ID)) +
    geom_col() + theme_classic(base_size = 20) + coord_flip()
```

Expression of protein markers across main cell types
```{r}
Protein_Res <- cbind.data.frame(as.matrix(pbmc.necs$ADT.cell.types), t(as.matrix(pbmc.necs[['ADT']]@data)))
colnames(Protein_Res)[1] <- "cell.types"
Protein_Res <- reshape2::melt(Protein_Res)
#Protein_Res$cell.types <- factor(Protein_Res$cell.types, levels = c("NonCytotoxic T", "Cytotoxic T", "NK", "BC", "Monocyte", "DC", "Myeloid"))
Protein_Res$variable <- factor(Protein_Res$variable, levels = c("CD3D","CD4","CD8A","NCAM1","CD19","CD33","CD14","FCGR3A","CD274","PDCD1"))
#Protein_Res$cell.types <- factor(Protein_Res$cell.types, levels = cell.types)

#Protein_Res <- Protein_Res %>% filter(cell.types != "MGK" & cell.types != "CD4 T Cytotoxic" & cell.types != "T gamma delta" & cell.types != "mDC" & cell.types != "pDC")

DimPlot(pbmc.necs, group.by = "ADT.cell.types")

require(ggplot2)
ggplot2::ggplot(Protein_Res, aes(
  x = variable, y = value,
  fill = variable), color = "grey") +
  ggplot2::geom_boxplot() +
  ggplot2::facet_wrap(~cell.types, scales="fixed", nrow = 2) +
  ylab("expression levels") + xlab("protein markers")+
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```

Expression of protein markers across final cell types
```{r}
Protein_Res <- cbind.data.frame(as.matrix(pbmc.necs$ct.consensus), t(as.matrix(pbmc.necs[['ADT']]@data)))
colnames(Protein_Res)[1] <- "cell.types"
Protein_Res <- reshape2::melt(Protein_Res)
#Protein_Res$cell.types <- factor(Protein_Res$cell.types, levels = c("NonCytotoxic T", "Cytotoxic T", "NK", "BC", "Monocyte", "DC", "Myeloid"))
Protein_Res$variable <- factor(Protein_Res$variable, levels = c("CD3D","CD4","CD8A","NCAM1","CD19","CD33","CD14","FCGR3A","CD274","PDCD1"))
Protein_Res$cell.types <- factor(Protein_Res$cell.types, levels = cell.types)

#Protein_Res <- Protein_Res %>% filter(cell.types != "MGK" & cell.types != "CD4 T Cytotoxic" & cell.types != "T gamma delta" & cell.types != "mDC" & cell.types != "pDC")

DimPlot(pbmc.necs, group.by = "ct.consensus")

require(ggplot2)
ggplot2::ggplot(Protein_Res, aes(
  x = variable, y = value,
  fill = variable), color = "grey") +
  ggplot2::geom_boxplot() +
  ggplot2::facet_wrap(~cell.types, scales="fixed", nrow = 2) +
  ylab("expression levels") + xlab("protein markers")+
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

Myeloid: average absolute scaled expression scores for each subtype
```{r}
Idents(pbmc.necs) <- "ct.consensus"
Myeloid <- subset(pbmc.necs, idents = c("M14","M16","mDC", "pDC", "MGK"))
Myeloid$ct.consensus <- droplevels(Myeloid$ct.consensus)


Myeloid.genesets <- c(immune.genesets[c("Monocyte","Mono.Mac","DC1","pDC.precursor","Megakaryocyte")])
Myeloid1 <- AddModuleScore(Myeloid, features = Myeloid.genesets, name = names(Myeloid.genesets))

cell.types <- levels(Myeloid1$ct.consensus)
plot.sets <- lapply(1:length(cell.types), function(x){
  Idents(Myeloid1) <- "ct.consensus"
  Myeloid1 <- subset(Myeloid1, idents = c(cell.types[x]))
  group <- Myeloid1$ct.consensus
  meta.data <- abs(Myeloid1@meta.data[c(16:20)])
  plot.sets <- vlnplot(Myeloid1, meta.data,group = group,cell.type = cell.types[x]) 
  
  return(plot.sets)
})
plot.sets

#gridExtra::grid.arrange(plot.sets[[1]],plot.sets[[2]], nrow = 1)
```


Lymphocytes: average absolute scaled expression scores for each subtype
```{r}
Idents(pbmc.necs) <- "ct.consensus"
Tcells <- subset(pbmc.necs, idents = c("CD4 T Naive", "CD4 T Memory","CD8 T","NK","BC Naive", "BC Memory"))
Tcells$ct.consensus <- droplevels(Tcells$ct.consensus)

CD4Memory.sig <- list("T cells CD4 memory" = unique(do.call(c,immune.genesets[c("T cells CD4 memory resting", "T cells CD4 memory activated")])))
NK.sig <- list("NK cells" = unique(do.call(c,immune.genesets[c("NK cells resting", "NK cells activated")])))

lymp.genesets <- c(immune.genesets[c("T cells CD4 naive")], CD4Memory.sig,immune.genesets[c("T cells CD8", "T cells gamma delta")], immune.genesets[c("B cells naive", "B cells memory")],NK.sig)
Tcells.genesets <- c(lymp.genesets)
Tcells1 <- AddModuleScore(Tcells, features = Tcells.genesets, name = names(Tcells.genesets))

cell.types <- levels(Tcells1$ct.consensus)
plot.sets <- lapply(1:length(cell.types), function(x){
  Idents(Tcells1) <- "ct.consensus"
  Tcells1 <- subset(Tcells1, idents = c(cell.types[x]))
  group <- Tcells1$ct.consensus
  meta.data <- abs(Tcells1@meta.data[c(16:22)])
  plot.sets <- vlnplot(Tcells1, meta.data,group = group,cell.type = cell.types[x]) 
  
  return(plot.sets)
})
plot.sets

#gridExtra::grid.arrange(plot.sets[[1]],plot.sets[[2]],plot.sets[[3]], plot.sets[[4]], plot.sets[[5]], plot.sets[[6]], nrow = 2)
```

Average expression of cytotoxic genes across lymphocyte populations
```{r}
DotPlot(pbmc.necs, features = c("GZMH","GZMB","GZMA","PRF1"), group.by = "ct.consensus")

```

Protein expression vs mRNA expression markers
```{r, eval = F, echo = F}
#centered log normalization of protein expression

DefaultAssay(pbmc.necs) <- "ADT"
pbmc.necs <- NormalizeData(pbmc.necs, assay = "ADT", normalization.method = "CLR") %>%
  ScaleData(assay = "ADT", split.by = "batch")

#All samples together
protein.markers <- rownames(pbmc.necs)
protein.feature <- FeaturePlot(pbmc.necs, features = rownames(pbmc.necs), cols = c("#f0f0f0", brewer.pal(9,"OrRd")), combine = F, max.cutoff = "q99", pt.size = 0.5)

protein.feature


#mRNA expression for all samples
#All samples together
DefaultAssay(pbmc.necs) <- "RNA"
RNA.feature <- FeaturePlot(pbmc.necs, features = protein.markers, cols = c("#f0f0f0", brewer.pal(9,"OrRd")), combine = F, max.cutoff = "q99", pt.size = 0.5)
RNA.feature

#combined protein and mRNA umap plots
umap.feature <- c(RNA.feature[1:5], protein.feature[1:5], RNA.feature[6:10], protein.feature[6:10])
pi <- umap.feature
umap.feature <- Reduce( `+`, pi ) +
      patchwork::plot_layout(ncol = 5)

#save umap of protein expression of all samples
pdf(paste0(work.dir, "/multimodal.umap.pdf"),20,15)
umap.feature
dev.off()

#boxplots of ADT markers of EL vs controls
p1 <- VlnPlot(pbmc.necs, features = protein.markers, assay = "ADT", slot = "data", same.y.lims = T, group.by = "age.group", combine = F)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
age.group.col <- c(cbPalette[6], cbPalette[8])

lapply(1:length(p1), function(x) {
  dat <- p1[[x]]$data
  p2 <- ggplot2::ggplot(dat, aes(ident, dat[,1])) + 
    geom_boxplot(aes(fill = ident)) +
    theme_classic(base_size = 20) +
    scale_fill_manual(values = age.group.col) +
    ylab("Expression Level") + xlab("age group") + ggtitle(colnames(dat)[1])
}
)

Idents(pbmc.necs) <- "age.group"
pbmc.CT <- subset(pbmc.necs, idents = "CT")
protein.markers <- c("CD3D","CD4","CD8A","NCAM1","CD19","CD33","CD14","FCGR3A","CD274","PDCD1")
pbmc.CT$ct.consensus <- factor(pbmc.CT$ct.consensus, levels = rev(cell.types))

p1 <- DotPlot(pbmc.CT, features = protein.markers, group.by = "ct.consensus", cols = c("#f0f0f0",  "#7F0000"), col.min = 0 )  + scale_size(limits = c(0,100), range = c(1,6)) 

pdf(paste0(work.dir, "/DotPlot.Protein.CT.pdf"),15,8)
p1
dev.off()

Idents(pbmc.necs) <- "age.group"
pbmc.EL <- subset(pbmc.necs, idents = "EL")
pbmc.EL$ct.consensus <- factor(pbmc.EL$ct.consensus, levels = rev(cell.types))


p2 <- DotPlot(pbmc.EL, features = protein.markers, group.by = "ct.consensus", cols = c("#f0f0f0",  "#7F0000"), col.min = 0) + scale_size(limits = c(0,100), range = c(1,6)) 

pdf(paste0(work.dir, "/DotPlot.Protein.EL.pdf"),15,8)
p2
dev.off()
```


CD274 and PDCD1
```{r, eval = F, echo = F}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
age.group.col <- c(cbPalette[6], cbPalette[8])

p1 <- VlnPlot(pbmc.necs, features = "PDCD1", group.by = "age.group", split.by = "ct.consensus", assay = "ADT", slot = "data") 
require(ggpubr)
cyt.T <- c("CD4 T Cytotoxic", "CD8 T", "T gamma delta")
p1 <- lapply(1:length(cyt.T), function(x){ggplot(p1$data[p1$data$split == cyt.T[x],], aes(ident,PDCD1, fill = ident))+ geom_boxplot() + facet_wrap(~split) +  scale_fill_manual(values = age.group.col) + theme_classic(base_size = 20) + stat_compare_means(method = "t.test")  
})


PDCD1.exp <- Reduce( `+`, p1 ) +
      patchwork::plot_layout(ncol = 3)

pdf(paste0(work.dir, "/Boxplot.CytT.PDCD1.pdf"),20,6)
PDCD1.exp
dev.off()

```

CITE-seq FACS-like Density Plots
```{r, eval = F, echo = F}
p1 <- FeatureScatter(pbmc.necs, "CD19", "CD3D", group.by = "ct.consensus", slot = "scale.data")

et_density <- function(x, y, ...) {
    dens <- MASS::kde2d(x, y, ...)
    ix <- findInterval(x, dens$x)
    iy <- findInterval(y, dens$y)
    ii <- cbind(ix, iy)
    return(dens$z[ii])
}
# tic()
d <- et_density(p1$data$CD19, p1$data$CD3D)
p1$data$density <- d

ggplot(p1$data, aes(CD19, CD3D, color = density)) +
        geom_point(size = 1) +
        scale_color_gradientn(colours = rev(brewer.pal(11, "Spectral")),
            trans = "sqrt") +
        theme_classic()
#Gating cells
cells.locate <- CellSelector(p1)
length(cells.locate)/ncol(pbmc.necs)
#Selecting T cells: 0.5225247 52%
#Selecting B cells: 0.02485903 2.5%

p2 <- FeatureScatter(pbmc.necs, "CD8A", "CD4", cells = cells.locate, group.by = "ct.consensus", slot = "scale.data")

d <- et_density(p2$data$CD8A, p2$data$CD4)
p2$data$density <- d

ggplot(p2$data, aes(CD8A, CD4, color = density)) +
        geom_point(size = 1) +
        scale_color_gradientn(colours = rev(brewer.pal(11, "Spectral")),
            trans = "sqrt") +
        theme_classic()

#Gating cells
CD4cells.locate <- CellSelector(p2)
length(CD4cells.locate)/length(cells.locate) 
#CD4 T cells: 0.5378278
CD8cells.locate <- CellSelector(p2)
length(CD8cells.locate)/length(cells.locate) 
#CD8 T cells: 0.3412625
```
