---
title: "Single cell analysis of CITE-seq data of centenarians from NECS Study: Integrating batches and identifying main cell types"
author: "Tanya Karagiannis, Stefano Monti, Paola Sebastiani"
output:
  html_document:
    theme: united
    toc: yes
  html_notebook:
    toc: yes
---

```{r}
knitr::opts_chunk$set(fig.width=12, fig.height=8) 

```

# Purpose: Identify main cell types based on protein expression
- Integrate Batch1 and Batch2 of NECS cohort
- Identify main cell types based on 10 protein markers

```{r}
library(dplyr)
library(Seurat)
library(useful)
library(cowplot)
library(SummarizedExperiment)
library(SingleR)
library(BiocParallel)
library(harmony, lib.loc = "/restricted/projectnb/montilab-p/R_pkgs/CentOS7/R_4.0.0/")

work.dir <- "/restricted/projectnb/montilab-p/projects/longevity/scCentenarians/scCentenarians/NECS_Analysis/"

Batch1 <- readRDS(file = paste0(work.dir, "Batch1/pbmc.cohort.Batch1.rds"))
DefaultAssay(Batch1) <- "RNA"

Batch2 <- readRDS(file = paste0(work.dir, "Batch2/pbmc.cohort.batch2.rds"))

```

## Change gene format
```{r}
Batch1.dat <- GetAssay(Batch1)
Batch1.counts <- Batch1.dat@counts
Batch1.log <- Batch1.dat@data
rownames(Batch1.counts) <- make.names(rownames(Batch1.counts))
rownames(Batch1.log) <- make.names(rownames(Batch1.log))

Batch1[['RNA']]@counts <- Batch1.counts
Batch1[['RNA']]@data <- Batch1.log

Batch2.dat <- GetAssay(Batch2)
Batch2.counts <- Batch2.dat@counts
Batch2.log <- Batch2.dat@data
rownames(Batch2.counts) <- make.names(rownames(Batch2.counts))
rownames(Batch2.log) <- make.names(rownames(Batch2.log))

Batch2[['RNA']]@counts <- Batch2.counts
Batch2[['RNA']]@data <- Batch2.log


```

## Merge samples, normalization, and metadata
```{r}
pbmc.necs <- merge(Batch1, Batch2, add.cell.ids = c("B1", "B2"), project = "GMcohort.combine") %>%
    Seurat::NormalizeData(verbose = FALSE) %>%
    FindVariableFeatures(selection.method = "vst", nfeatures = 2000) 
    #Seurat::SCTransform(vars.to.regress = c("batch", "percent.mt"), verbose = FALSE)

head(pbmc.necs@meta.data)

meta.data <- pbmc.necs@meta.data[,c("orig.ident","nCount_RNA","nFeature_RNA","nCount_ADT","nFeature_ADT","sample.ID","age.group","age","sex","percent.mt","batch", "ct.consensus")]

pbmc.necs@meta.data <- meta.data

pbmc.necs@meta.data$batch <- c(rep("COHORT.B1", nrow(Batch1@meta.data)), rep("COHORT.B2", nrow(Batch2@meta.data)))
pbmc.necs <- ScaleData(pbmc.necs, vars.to.regress = "percent.mt")
```

## PCA Analysis
```{r}
pbmc.necs <- pbmc.necs %>% RunPCA(features = VariableFeatures(pbmc.necs), npcs = 50, verbose = FALSE)
p1 <- DimPlot(object = pbmc.necs, reduction = "pca", pt.size = .1, group.by = "batch")
p2 <- VlnPlot(object = pbmc.necs, features = "PC_1", group.by = "batch", pt.size = .1)
plot_grid(p1,p2)

```

## Batch correction of samples in PCA space 
- Harmony (Korunsky et al 2019 Nature Methods)
```{r harmony correction}
pbmc.necs <- pbmc.necs %>% 
    RunHarmony(group.by.vars ="batch", plot_convergence = TRUE)

#saveRDS(pbmc.necs, file = "./pbmc.necs.rds")

p1 <- DimPlot(object = pbmc.necs, reduction = "harmony", pt.size = .1, group.by = "batch")
p2 <- VlnPlot(object = pbmc.necs, features = "harmony_1", group.by = "batch", pt.size = .1)
plot_grid(p1,p2)

```

## Run UMAP and clustering based on top 20 Harmony-corrected PCs
```{r harmony umap}
set.seed(822)
pbmc.necs <- pbmc.necs %>% 
    RunUMAP(reduction = "harmony", dims = 1:20) %>%
    FindNeighbors(reduction = "harmony", dims = 1:20) %>% 
    FindClusters(resolution = 0.8) %>% 
    identity()
```

## Visualize UMAP and clusters
```{r}
p1 <- DimPlot(object = pbmc.necs, reduction = "umap", pt.size = .1, group.by = "seurat_clusters", label = TRUE)
p2 <- DimPlot(object = pbmc.necs, reduction = "umap", pt.size = .1, group.by = "batch")
plot_grid(p1,p2)
```

## Protein expression levels to identify main cell types
- 10 protein markers
- normalize and scale protein expression data
```{r}
DefaultAssay(pbmc.necs) <- "ADT"
pbmc.necs <- NormalizeData(pbmc.necs, assay = "ADT", normalization.method = "CLR") %>%
  ScaleData(assay = "ADT")
```

## Expression of proteins per cluster
```{r}
Res <- cbind.data.frame(as.matrix(pbmc.necs$seurat_clusters), t(as.matrix(pbmc.necs[['ADT']]@data)))
colnames(Res)[1] <- "clusters"
ResFlat <- reshape2::melt(Res)
colnames(ResFlat) <- c("clusters","protein.markers", "expression")

ResFlat$protein.markers <- factor(ResFlat$protein.markers, levels = c("CD3D","CD4","CD8A","NCAM1","CD19","CD33","CD14","FCGR3A", "CD274","PDCD1"))
ResFlat$clusters <- factor(ResFlat$clusters, levels = 0:17)

ggplot2::ggplot(ResFlat, aes(
  x = protein.markers, y = expression,
  fill = protein.markers)) +
  ggplot2::geom_boxplot() +
  ggplot2::facet_wrap(~clusters, scales="fixed") +
  theme_classic(base_size = 20) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Select main cell types based on protein levels
```{r}
#CD4 T: Cluster 2, 3, 16
#CD8 T: Cluster 1, 6, 12 
#CD4 + CD8 T Mix: Cluster 7
#NK cells: Cluster 0, 10
#B cells: Cluster 8
#CD14 Mono: Cluster 4, 5, 9, 15
#CD16 Mono: Cluster 11
#Myeloid: 13
#CD4 Pos Myeloid (pDC): 17
#CD274+: 14

#Set main cell types based on protein expression of the 10 cell surface markers for each cluster
Idents(pbmc.necs) <- "seurat_clusters"
new.cluster.ids <- c("NK","CD8 T", "CD4 T", "CD4 T", "M14", "M14", "CD8 T", "CD4 + CD8 T", "B", "M14","NK","M16","CD8 T","Myeloid", "CD274+","M14", "CD4 T","CD4+ Myeloid")
#new.cluster.ids <- c("CD8 T","NK", "M14", "CD4 T", "CD4 T", "M14", "CD4 T", "CD8 T", "CD4 T Cytotoxic", "B","M14","NK","M16","T gamma delta", "CD4 T","Myeloid", "Myeloid","M14","CD4 T","CD4 pos")


names(new.cluster.ids) <- levels(pbmc.necs)
pbmc.necs <- RenameIdents(pbmc.necs, new.cluster.ids)
pbmc.necs$ADT.cell.types <- Idents(pbmc.necs)

#saveRDS(pbmc.necs, file = paste0(work.dir, "Data/pbmc.necs.rds"))
```

## Visualize UMAP based on ADT-based cell types and batch
```{r}
p1 <- DimPlot(object = pbmc.necs, reduction = "umap", pt.size = .1, group.by = "ADT.cell.types", label = TRUE)
p2 <- DimPlot(object = pbmc.necs, reduction = "umap", pt.size = .1, group.by = "batch")
plot_grid(p1,p2)
```



# SingleR Analysis 
Comparison of main cell types from public available datasets with protein-based main cell types in NECS dataset

## Load public datasets and NECS data
```{r, eval = F}
pbmc.natgen <- readRDS("/restricted/projectnb/montilab-p/projects/longevity/scCentenarians/scCentenarians/PublicData_Reanalysis/natgen2018/data/natgen2018.rds")
pbmc.pnas <- readRDS("/restricted/projectnb/montilab-p/projects/longevity/scCentenarians/scCentenarians/PublicData_Reanalysis/pnas2019/data/pnas2019.rds")
pbmc.necs <- readRDS("/restricted/projectnb/montilab-p/projects/longevity/scCentenarian/NECS_Analysis/Data/pbmc.necs.updated.rds")
DefaultAssay(pbmc.necs) <- "RNA"

pbmc.natgen <- SummarizedExperiment(assays=list(logcounts=pbmc.natgen[['RNA']]@data), colData=pbmc.natgen@meta.data)
pbmc.pnas <- SummarizedExperiment(assays=list(logcounts=pbmc.pnas[['RNA']]@data), colData=pbmc.pnas@meta.data)
pbmc.necs <- SummarizedExperiment(assays=list(logcounts=pbmc.necs[['RNA']]@data), colData=pbmc.necs@meta.data)

```

## SingleR classification with DE genes (one vs all result)
```{r, eval = F}
class.res <- SingleR(test = pbmc.necs,
    ref = list(natgen=pbmc.natgen, pnas=pbmc.pnas), 
    labels = list(pbmc.natgen$ct.consensus, pbmc.pnas$ct.consensus),
    method = "cluster",
    clusters = pbmc.necs$seurat_clusters)
    #BPPARAM = MulticoreParam(workers=8))

#saveRDS(class.res, file = "Data/class.res.rds")

table(class.res$labels)
table(class.res$reference)  

```

## Visualize comparison with protein-based cell types identified
```{r, eval = F}
plotScoreHeatmap(class.res)
Idents(pbmc.necs) <- "seurat_clusters"
new.cluster.ids <- class.res$labels
names(new.cluster.ids) <- levels(pbmc.necs)
pbmc.necs <- RenameIdents(pbmc.necs, new.cluster.ids)
pbmc.necs$singleR.labels <- Idents(pbmc.necs)
knitr::kable(table(pbmc.necs$ADT.cell.types, pbmc.necs$singleR.labels))

set.mat <- table(pbmc.necs$ADT.cell.types, pbmc.necs$singleR.labels)
set.mat

#98.8% accuracy in detection
# library(caret) 
# confusionMatrix(set.mat)

p1 <- DimPlot(pbmc.necs, group.by = "ADT.cell.types")
p2 <- DimPlot(pbmc.necs, group.by = "singleR.labels")
plot_grid(p1, p2)

#saveRDS(pbmc.necs, file = paste0(work.dir,"Data/pbmc.necs.rds"))
```









